### 2020-07-22 PCM_preprocessing_final_20200715.Rmd
pre-processing

---
title: "PCM data analysis"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 12
    fig_height: 6
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,autodeep=TRUE,cache.comments=FALSE,warning=FALSE,message=FALSE,cache.lazy=FALSE)
```

```{r,echo=FALSE}
library(openxlsx)
library(RPlotRNAseq)
library(SCISSOR)
library(lattice)
library(quantreg)
library(ggplot2)
library(pscl)
library(boot)
library(MASS)
library(dplyr)

# today = Sys.Date()
dataDir = "./rawData/"
```

# Data files

*_Raw data files._*

* PatientVars_zip.csv   
* R21_Sample_wDxDate.xlsx        
* joined_zipcode.xlsx    
* Copy.xlsx (Occupation, Primary payer, marital status)      
* pcm.xlsx

*_Manipulated data files._*

* PatientVars_zip.csv -> PatientVars_zip_neededVars.xlsx   
    - Only needed variables (profile_key, dx_date_use, Marital_Status_at_DX_Desc, ssn_wcc_num, zip, accession_no, state, race_NIH, race_simple, stage_merged, stage_simple, gender, age_at_dx, age_binary, first_chemo, first_surgery) were selected.    
    - first_chemo & first_surgery variables were reformatted as "Date" to be acceptable by R.

* R21_Sample_wDxDate.xlsx -> newDxDate.xlsx    
    - Cleaned datasets with new dx_date (initial diagnosis date) was released (3/12/2020) - R21_Sample_wDxDate.    
    - newDxDate file is a copy of R21_Sample_wDxDate file to modify the original variables for the downstream analysis. dx_date was renamed to dx_date_use (which is the same as in PatientVars_zip3.xlsx).   
    - All patients in PatientVars_zip3 file are also recorded in newDxDate file.    
    - 174 patients are missing in dx_date_use.

* PatientVars_zip3.xlsx + newDxDate.xlsx -> patients_file_new.xlsx

* joined_zipcode.xlsx -> poverty_file_new.xlsx   
    - Poverty, vacancy, walkability, and unemployment were taken as group mean by zip

* Re-merge the datasets based on the new dx_date_use (newDxDate file).

original dx_date_use has been replaced by the new dx_date_use.

## In PatientVars_zip3, replace the dx_date_use variable by the new dx_date_use in newDxDate.xlsx.

```{r,echo=FALSE,eval=FALSE}
patients_file = read.xlsx(xlsxFile=paste0(dataDir,"PatientVars_zip_neededVars.xlsx"),sheet=1,detectDates=TRUE)
patients_file$ssn_wcc_num=as.character(patients_file$ssn_wcc_num)
# dim(patients_file)
# head(patients_file)
# length(unique(patients_file$profile_key))

newDxDate_file = read.xlsx(xlsxFile=paste0(dataDir,"newDxDate.xlsx"),sheet=1,detectDates=TRUE)
# dim(newDxDate_file)
# head(newDxDate_file)
# length(unique(newDxDate_file$profile_key))

## Collapse with the newDxDate and PatientVars_zip3 -> patients_file2
patients_file2 = merge(x=patients_file,y=newDxDate_file,by="profile_key")
# head(patients_file2)
# dim(patients_file2)

cat(paste("Number of different dx_date_use =",length(which(! patients_file2$dx_date_use.x==patients_file2$dx_date_use.y))),"\n")
cat(paste("Number of NA's in dx_date_use =",length(which(is.na(patients_file2$dx_date_use.y)))),"\n")

## Make new PatientVars file. 
# patients_file3 = patients_file2[,c(1,15,3,19,5:10,30,12:14,22:29)]
patients_nvars = dim(patients_file)[2]
patients_file3 = patients_file2[,c(1,(patients_nvars+1),3:patients_nvars)] # Should be carefully chosen!
# head(patients_file3)
colnames(patients_file3)[2] = "dx_date_use"
colnames(patients_file3)[4] = "ssn_wcc_num"

# first_chemo & first_surgery variable class - should be fixed for strange dates
class(patients_file3$first_chemo)
class(patients_file3$first_surgery)
patients_file3$first_chemo = as.Date(patients_file3$first_chemo)
patients_file3$first_surgery = as.Date(patients_file3$first_surgery)
class(patients_file3$first_chemo)
class(patients_file3$first_surgery)
patients_file3$first_chemo[which(grepl(pattern="[****]",patients_file3$first_chemo)==T)] = NA
patients_file3$first_surgery[which(grepl(pattern="[****]",patients_file3$first_surgery)==T)] = NA

## Final dataset
patients_file = patients_file3
head(patients_file)
# write.xlsx(x=patients_file,file=paste0(dataDir,"patients_file_new.xlsx"))
```

## Poverty data

* "joined_zipcode" file merged to get data about poverty, vacancy, unemployment levels for more zip codes.   
    - The main difference from the 'PCM_preprocessing' file is that the zip_poverty file (POVERTY, VACANCY, WALKABILITY, UNEMPLOYMENT) is replaced by "joined_zipcode" file.   
    - All zip codes in "zip_poverty" file are included in "joined_zipcode" file.   
    - And the statistics are strongly matched to each other.    
    - Good for the replacement.

```{r,echo=FALSE,eval=FALSE}
## New "joined_zipcode" file
joined_zipcode_file=read.xlsx(xlsxFile=paste0(dataDir,"joined_zipcode.xlsx"),sheet=1,colNames=T,
                              skipEmptyCols=F,skipEmptyRows=F)
# dim(joined_zipcode_file)
# head(joined_zipcode_file)
# sapply(joined_zipcode_file,class)
joined_zipcode_file$zip = as.character(joined_zipcode_file$zip)

# head(joined_zipcode_file[,c(2,4,11,12,13,14,27)])
joined_zipcode = joined_zipcode_file[,c(2,4,11,12,13,14,27)]

# table(joined_zipcode$zip)
joined_poverty_mean=aggregate(. ~ zip, joined_zipcode, mean)
colnames(joined_poverty_mean) = c("ZIP_CODE_1","Population2010","CANCER","POVERTY","VACANCY","WALKABILITY","UNEMPLOYMENT")

poverty_mean=joined_poverty_mean[,c(1,4:7)]
head(poverty_mean)
dim(poverty_mean)
# write.xlsx(x=poverty_mean,file=paste0(dataDir,"poverty_file_new.xlsx"))
```

* Poverty data (from "joined_zipcode")

    - For poverty variable, group mean used for each zip code. -> poverty_file_new.xlsx        
    - Only 43 zip codes are documented in poverty file. On the other hand, there are 165 zip codes (TN) in Patients file. This will produce too many missing values for the poverty variable.

# Merge datasets

```{r,echo=FALSE}
patients_file = read.xlsx(xlsxFile=paste0(dataDir,"patients_file_new.xlsx"),sheet=1,detectDates=TRUE)
class(patients_file$first_chemo)
patients_file$first_chemo = as.Date(patients_file$first_chemo)
class(patients_file$first_chemo)

poverty_file = read.xlsx(xlsxFile=paste0(dataDir,"poverty_file_new.xlsx"),sheet=1,detectDates=TRUE)

copy_file1=read.xlsx(xlsxFile=paste0(dataDir,"Copy.xlsx"),sheet=1,detectDates=TRUE)
copy_file2=read.xlsx(xlsxFile=paste0(dataDir,"Copy.xlsx"),sheet=2,detectDates=TRUE)
copy_file3=read.xlsx(xlsxFile=paste0(dataDir,"Copy.xlsx"),sheet=3,detectDates=TRUE)
# copy_file=rbind(copy_file1,copy_file2,copy_file3)[,c(1,5,13,17,18,23,24,25,26,27,57,58)]

copy_file0 = rbind(copy_file1,copy_file2,copy_file3)
copy_file <-
  copy_file0 %>%
  select("Accession.Number","Social.Security.Number","Vital.Status-Desc",
         "Text.-.Usual.Occupation","Text.-.Usual.Industry",
         "Date.of.Initial.Diagnosis","Age.at.Diagnosis","Primary.Payer.at.DX-Desc",
         "Marital.Status.at.DX-Desc","Tobacco.History-Desc",
         "Date.of.First.Surgery","Cancer.Status-Desc","Date.1st.Recurrence",
         "Postal.Code.at.Diagnosis","State.at.Diagnosis-Desc",
         "Primary.Site-Desc","Histo/Behavior.ICD-O-3-Desc","Tumor.Size-MM",
         "ER","PR","HER2","RT.Site-Desc","Type.1st.Recurrence-Desc")

copy_file$Social.Security.Number=as.character(as.numeric(copy_file$Social.Security.Number))

pcm_file=read.xlsx(xlsxFile=paste0(dataDir,"pcm.xlsx"),sheet=1,detectDates=TRUE)

pcm_file[which(pcm_file$profile_key=="1665"),]
patients_file[which(patients_file$profile_key=="1665"),]
```

* The cohorts of PCM file and Patients file are equal.

```{r,echo=FALSE}
pcm_profile_key=sort(unique(pcm_file$profile_key))
patients_profile_key=sort(unique(patients_file$profile_key))
N = length(patients_profile_key)

length(which(! pcm_profile_key %in% patients_profile_key))
length(which(! patients_profile_key %in% pcm_profile_key))
cat(paste("\t Total number of samples =",N),"\n")
```

```{r,echo=FALSE}
profile_key = patients_profile_key
## PatientVars_zip file
sortbydate=function(x) {
  temp_rows=which(patients_file$profile_key==x)
  if (sum(is.na(patients_file$dx_date_use[temp_rows]))==length(temp_rows)) {
    temp_rows[1]
  } else {
    temp_rows[which(patients_file$dx_date_use[temp_rows]==min(patients_file$dx_date_use[temp_rows],na.rm=TRUE))][1]
  }
}
patients_rows=apply(matrix(profile_key,ncol=1),1,FUN=sortbydate)
patients_sort=patients_file[patients_rows,]
# head(patients_sort)
# tail(patients_sort)
cat(paste("\t Number of NA's in dx_date_use in PatientsVar =",sum(is.na(patients_sort$dx_date_use))),"\n")

```

patients_file + copy_file + poverty_file

* Merge patients file and copy file by SSN and then merge it with poverty file by ZIP

```{r,echo=FALSE}
m1 = merge(x=patients_sort,y=copy_file,by.x="ssn_wcc_num",by.y="Social.Security.Number",all=T)
dim(m1)
m1 = m1[order(m1$profile_key),] # Sort by profile key
m2 = m1[which(m1$ssn_wcc_num %in% patients_sort$ssn_wcc_num),]
dim(m2)
mergedData = merge(m1,poverty_file,by.x="zip",by.y="ZIP_CODE_1",all=T)
mergedData = mergedData[order(mergedData$profile_key),]
dim(mergedData)
# mergedData=merge(pcm_sort,m2,by="profile_key",all=T)
```

```{r,echo=FALSE,eval=FALSE}
length(unique(copy_file$Social.Security.Number))

length(unique(m1$profile_key))
length(unique(mergedData$zip))
length(unique(mergedData$profile_key))
dim(mergedData)
```

## Merging datasets naively gives many duplicates and this should be addressed.

1. Delete the perfect duplicates, i.e. the rows with the exactly same information.  (Number of rows: 11025 $\rightarrow$ 7689)

N = 3836

```{r}
mergedData_uniq1 = distinct(mergedData)
dim(mergedData_uniq1)
length(unique(mergedData_uniq1$profile_key))
```

2. Remove redundant samples. 
    - States other than TN (7714 $\rightarrow$ 2673)      
    - Profile key: (3836 -> 2628)

```{r,echo=FALSE}
cat(paste("All distinct states =",paste(unique(mergedData_uniq1$state),collapse=",")),"\n")
```

```{r}
mergedData_uniq2 = mergedData_uniq1[which(mergedData_uniq1$state=="TN"),]
dim(mergedData_uniq2)
length(unique(mergedData_uniq2$profile_key))
length(unique(mergedData_uniq2$zip))
```

3. Other duplicates using profile_key

    - There are 43 duplicated profile keys in the merged data set.     
    - e.g. profile key==90006

```{r,echo=FALSE}
dup_samples=which(duplicated(mergedData_uniq2$profile_key)==1)
length(mergedData_uniq2$profile_key[dup_samples])
length(unique(mergedData_uniq2$profile_key[dup_samples]))

mergedData_uniq2[which(mergedData_uniq2$profile_key==mergedData_uniq2$profile_key[dup_samples[1]]),]

cat(paste("\t ",length(dup_samples)," duplicated samples =",paste(mergedData_uniq2$profile_key[dup_samples],collapse=",")),"\n")
```

4. Remove the 45 duplicates (2628 -> 2584)

```{r}
profile.split = split(x=which(mergedData_uniq2$profile_key %in% mergedData_uniq2$profile_key[dup_samples]),
                      f=mergedData_uniq2$profile_key[which(mergedData_uniq2$profile_key %in% mergedData_uniq2$profile_key[dup_samples])])

discard_dup = function(x) {
  x[-which.min(apply(mergedData_uniq2[x,],1,function(x) length(which(is.na(x)==T))))]
}
discard_ids = do.call("c",sapply(profile.split,discard_dup))
length(discard_ids)

mergedData_uniq3 = mergedData_uniq2[-discard_ids,]

dim(mergedData_uniq3)
length(unique(mergedData_uniq3$profile_key))
length(unique(mergedData_uniq3$zip))

```

```{r,echo=FALSE}
N=length(unique(mergedData_uniq3$profile_key))
cat(paste("\t Total number of unique samples in the merged data (N) =",N),"\n")
```

```{r,eval=FALSE}
# today = "2020-07-15"
# write.xlsx(x=mergedData_uniq3,file=paste0(dataDir,"PatientsData_merged_",today,"_raw.xlsx"))
```

## Further treatment for PatientsData file.

* Variable names changed   
* Remove duplicated variable    
* missing -> unknown   
* data structure     
    - NA -> Unknown
    - experimental variables: 
    - control variables: age, stage, primary site,

```{r,echo=FALSE}
merged_file=read.xlsx(xlsxFile=paste0(dataDir,"PatientsData_merged_2020-07-15_raw.xlsx"),sheet=1,detectDates=TRUE)

merged_file$zip = sapply(as.character(merged_file$zip),FUN=function(x){paste(strsplit(x,split="")[[1]][1:5],collapse="")})

colnames(merged_file)[which(colnames(merged_file)=="Marital_Status_at_DX_Desc")] = "Marital_Status_at_DX"
colnames(merged_file)[which(colnames(merged_file)=="Vital.Status-Desc")] = "Vital.Status"
colnames(merged_file)[which(colnames(merged_file)=="Text.-.Usual.Occupation")] = "Usual.Occupation"
colnames(merged_file)[which(colnames(merged_file)=="Text.-.Usual.Industry")] = "Usual.Industry"
colnames(merged_file)[which(colnames(merged_file)=="Primary.Site-Desc")] = "Primary.Site"
colnames(merged_file)[which(colnames(merged_file)=="Primary.Payer.at.DX-Desc")] = "Primary.Payer.at.DX"
colnames(merged_file)[which(colnames(merged_file)=="Marital.Status.at.DX-Desc")] = "Marital.Status.at.DX"
colnames(merged_file)[which(colnames(merged_file)=="Tobacco.History-Desc")] = "Tobacco.History"
colnames(merged_file)[which(colnames(merged_file)=="Tumor.Size-MM")] = "Tumor.Size.MM"
colnames(merged_file)[which(colnames(merged_file)=="Cancer.Status-Desc")] = "Cancer.Status"
colnames(merged_file)[which(colnames(merged_file)=="State.at.Diagnosis-Desc")] = "State.at.Diagnosis"
colnames(merged_file)[which(colnames(merged_file)=="RT.Site-Desc")] = "RT.Site"
colnames(merged_file)[which(grepl(pattern="Type.1st",colnames(merged_file))==T)] = "Recurrence"
colnames(merged_file)[which(grepl(pattern="Behavior",colnames(merged_file))==T)] = "Histology"
colnames(merged_file)[which(grepl(pattern="metasta",colnames(merged_file))==T)] = "Metastasis"

merged_file = merged_file[,-which(colnames(merged_file)=="Marital.Status.at.DX")]

cat(paste("Number of missing values for each variable:"),"\n")
for (jc in 1:dim(merged_file)[2]) {
  cat(paste(jc,"|",colnames(merged_file)[jc],"|",length(which(is.na(merged_file[,jc])==1))),"\n")
}

## missing values
hist(apply(merged_file,1,FUN=function(x){sum(is.na(x))}))
# write.xlsx(x=merged_file[sample1,c(3,4,32:35)],file=paste0(dataDir,"PCM_merged_POVERTY_nomissing.xlsx"))
```

* Change the values for each variable to favoriable format:

```{r,echo=FALSE}
# Cancer status
merged_file$Cancer.Status[which(is.na(merged_file$Cancer.Status)==1)] = "Unknown"
merged_file$Cancer.Status[which(grepl(pattern="Unknown whether",x=merged_file$Cancer.Status)==1)] = "Unknown"
merged_file$Cancer.Status[which(grepl(pattern="No evidence",x=merged_file$Cancer.Status)==1)] = "No_evidence"
merged_file$Cancer.Status[which(grepl(pattern="Evidence of",x=merged_file$Cancer.Status)==1)] = "Evidence"
table(merged_file$Cancer.Status)

# Cancer.Status.n = rep(0,dim(merged_file)[1])
# Cancer.Status.n[which(merged_file$Cancer.Status=="Unknown")] = 2
# Cancer.Status.n[which(merged_file$Cancer.Status=="Evidence")] = 1
Cancer.Status.f = factor(merged_file$Cancer.Status)
Cancer.Status.f = relevel(Cancer.Status.f,"No_evidence")
table(Cancer.Status.f)

## Cancer stage
stage.f = factor(merged_file$stage_simple)
stage.f = relevel(stage.f,"1")
table(stage.f)

## Age
age.f = factor(merged_file$age_binary)
table(age.f)

## Race
table(merged_file$race_simple)
race.f = factor(merged_file$race_simple)
race.f = relevel(race.f,"White")
table(race.f)

## Marital status
merged_file$Marital_Status_at_DX[which(is.na(merged_file$Marital_Status_at_DX)==1)] = "Unknown"
merged_file$Marital_Status_at_DX[which(merged_file$Marital_Status_at_DX=="Married (including common")] = "Married"
merged_file$Marital_Status_at_DX[which(merged_file$Marital_Status_at_DX=="Single (never married)")] = "Single"
merged_file$Marital_Status_at_DX[which(merged_file$Marital_Status_at_DX=="Unmarried or Domestic Par")] = "Single"
merged_file$Marital_Status_at_DX[which(merged_file$Marital_Status_at_DX=="Separated")] = "Divorced"

marital.status.f = factor(merged_file$Marital_Status_at_DX)
marital.status.f = relevel(marital.status.f,"Married")
table(marital.status.f)

## Usual occupation
merged_file$Usual.Occupation[which(grepl(pattern="retired",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Retired"

merged_file$Usual.Occupation[which(grepl(pattern="disabled",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Disabled"

merged_file$Usual.Occupation[which(grepl(pattern="unemployed",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Unemployed"
merged_file$Usual.Occupation[which(grepl(pattern="not employed",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Unemployed"

merged_file$Usual.Occupation[which(grepl(pattern="self",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Self_employed"

merged_file$Usual.Occupation[which(grepl(pattern="unknown",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Unknown"
merged_file$Usual.Occupation[which(grepl(pattern="none",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Unknown"
merged_file$Usual.Occupation[which(grepl(pattern="Not listed",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Unknown"
merged_file$Usual.Occupation[which(grepl(pattern="Not employed",x=merged_file$Usual.Occupation,ignore.case=T)==1)] = "Unknown"

merged_file$Usual.Occupation[which(is.na(merged_file$Usual.Occupation)==1)] = "Unknown"

merged_file$Usual.Occupation[which(! merged_file$Usual.Occupation %in% c("Retired","Disabled","Unknown","Unemployed","Self-employed"))] = "Employed"

# length(which(! merged_file$Usual.Occupation %in% c("Retired","Disabled","Unknown","Unemployed","Self-employed")))

Usual.Occupation.f = factor(merged_file$Usual.Occupation)
Usual.Occupation.f = relevel(Usual.Occupation.f,"Employed")
table(Usual.Occupation.f)

## Primary Payer
merged_file$Primary.Payer.at.DX[which(grepl(pattern="tricare",x=merged_file$Primary.Payer.at.DX,ignore.case=T)==1)] = "Medicaid"
merged_file$Primary.Payer.at.DX[which(grepl(pattern="medicare",x=merged_file$Primary.Payer.at.DX,ignore.case=T)==1)] = "Medicaid"
merged_file$Primary.Payer.at.DX[which(grepl(pattern="medicaid",x=merged_file$Primary.Payer.at.DX,ignore.case=T)==1)] = "Medicaid"
merged_file$Primary.Payer.at.DX[which(grepl(pattern="private",x=merged_file$Primary.Payer.at.DX,ignore.case=T)==1)] = "Private"
merged_file$Primary.Payer.at.DX[which(grepl(pattern="unknown",x=merged_file$Primary.Payer.at.DX,ignore.case=T)==1)] = "Unknown"
merged_file$Primary.Payer.at.DX[which(grepl(pattern="not",x=merged_file$Primary.Payer.at.DX,ignore.case=T)==1)] = "Unknown"
merged_file$Primary.Payer.at.DX[which(is.na(merged_file$Primary.Payer.at.DX)==1)] = "Unknown"

merged_file$Primary.Payer.at.DX[which(! merged_file$Primary.Payer.at.DX %in% c("Medicaid","Unknown","Private"))] = "Insurance"

Primary.Payer.f = factor(merged_file$Primary.Payer.at.DX)
Primary.Payer.f = relevel(Primary.Payer.f,"Private")
table(Primary.Payer.f)

## Merge factor variables to the original data set
merged_file$Cancer.Status.f = Cancer.Status.f
merged_file$stage.f = stage.f
merged_file$age.f = age.f
merged_file$race.f = race.f
merged_file$marital.status.f = marital.status.f
merged_file$Usual.Occupation.f = Usual.Occupation.f
merged_file$Primary.Payer.f = Primary.Payer.f

table(merged_file$Tobacco.History)
merged_file$Tobacco.History[is.na(merged_file$Tobacco.History)] = "Unknown"
merged_file$Tobacco.History[grepl(pattern="Never",merged_file$Tobacco.History)] = "Never"
merged_file$Tobacco.History[grepl(pattern="Smoke|smoke",merged_file$Tobacco.History)] = "Smoked"
merged_file$Tobacco.History[grepl(pattern="positive",merged_file$Tobacco.History)] = "Smoked"
merged_file$Tobacco.History[grepl(pattern="Unknown",merged_file$Tobacco.History)] = "Unknown"
merged_file$Tobacco.History[grepl(pattern="    ",merged_file$Tobacco.History)] = "Unknown"
table(merged_file$Tobacco.History)
merged_file$Tobacco.History = factor(merged_file$Tobacco.History,
                             levels=c("Never","Smoked","Unknown"))

# table(merged_file$Histology)
merged_file$Histology[which(is.na(merged_file$Histology)==T)] = "Unknown"
merged_file$Histology[(grepl(pattern="duct",merged_file$Histology,ignore.case=T) & grepl(pattern="lob",merged_file$Histology,ignore.case=T))] = "ductal"
merged_file$Histology[grepl(pattern="duct",merged_file$Histology,ignore.case=T)] = "ductal"
merged_file$Histology[grepl(pattern="lob",merged_file$Histology,ignore.case=T)] = "lobular"
# table(merged_file$Histology[!grepl(pattern="lobular|ductal",merged_file$Histology,ignore.case=T)])
merged_file$Histology[!grepl(pattern="lobular|ductal|Unknown",merged_file$Histology,ignore.case=T)] = "Other"
table(merged_file$Histology)

table(merged_file$HER2)
merged_file$HER2[which(is.na(merged_file$HER2)==T)] = "Unknown"
merged_file$HER2[grepl(pattern="Negative",merged_file$HER2,ignore.case=T)] = "Negative.Normal"
merged_file$HER2[grepl(pattern="Positive",merged_file$HER2,ignore.case=T)] = "Positive.Elevated"
merged_file$HER2[grepl(pattern="not",merged_file$HER2,ignore.case=T)] = "Test.not.done"

table(merged_file$ER)
merged_file$ER[which(is.na(merged_file$ER)==T)] = "Unknown"
merged_file$ER[grepl(pattern="Negative",merged_file$ER,ignore.case=T)] = "Negative.Normal"
merged_file$ER[grepl(pattern="Positive",merged_file$ER,ignore.case=T)] = "Positive.Elevated"
merged_file$ER[grepl(pattern="not",merged_file$ER,ignore.case=T)] = "Test.not.done"

table(merged_file$PR)
merged_file$PR[which(is.na(merged_file$PR)==T)] = "Unknown"
merged_file$PR[grepl(pattern="Negative",merged_file$PR,ignore.case=T)] = "Negative.Normal"
merged_file$PR[grepl(pattern="Positive",merged_file$PR,ignore.case=T)] = "Positive.Elevated"
merged_file$PR[grepl(pattern="not",merged_file$PR,ignore.case=T)] = "Test.not.done"

table(merged_file$Recurrence)
merged_file$Recurrence[which(is.na(merged_file$Recurrence)==T)] = "Unknown"
merged_file$Recurrence[grepl(pattern="never been disease-free",merged_file$Recurrence,ignore.case=T)] = "Recurrence"
merged_file$Recurrence[grepl(pattern="ever disease-free",merged_file$Recurrence,ignore.case=T)] = "Recurrence"
merged_file$Recurrence[grepl(pattern="disease-free",merged_file$Recurrence,ignore.case=T)] = "disease-free"
merged_file$Recurrence[!grepl(pattern="disease-free|Unknown",merged_file$Recurrence,ignore.case=T)] = "Recurrence"
merged_file$Recurrence[grepl(pattern="unknown",merged_file$Recurrence,ignore.case=T)] = "Unknown"

merged_file$Primary.Site[which(is.na(merged_file$Primary.Site)==T)] = "Unknown"
merged_file$Primary.Site.1 = merged_file$Primary.Site.2 = merged_file$Primary.Site
merged_file$Primary.Site.1[grepl(pattern="lower",merged_file$Primary.Site.1,ignore.case=T)] = "Lower"
merged_file$Primary.Site.1[grepl(pattern="upper",merged_file$Primary.Site.1,ignore.case=T)] = "Upper"
merged_file$Primary.Site.1[grepl(pattern="central",merged_file$Primary.Site.1,ignore.case=T)] = "Central"
merged_file$Primary.Site.1[!grepl(pattern="upper|central|lower|unknown",merged_file$Primary.Site.1,ignore.case=T)] = "Unknown"

table(merged_file$Primary.Site.2)
merged_file$Primary.Site.2[grepl(pattern="inn",merged_file$Primary.Site.2,ignore.case=T)] = "Inner"
merged_file$Primary.Site.2[grepl(pattern="out",merged_file$Primary.Site.2,ignore.case=T)] = "Outer"
merged_file$Primary.Site.2[grepl(pattern="mid",merged_file$Primary.Site.2,ignore.case=T)] = "Midline"
merged_file$Primary.Site.2[grepl(pattern="over",merged_file$Primary.Site.2,ignore.case=T)] = "Overlapping"
merged_file$Primary.Site.2[!grepl(pattern="inner|midline|outer|unknown|overlapping",merged_file$Primary.Site.2,ignore.case=T)] = "Unknown"

table(merged_file$Metastasis)
merged_file$Metastasis[which(is.na(merged_file$Metastasis)==T)] = "No"
merged_file$Metastasis[which(merged_file$Metastasis=="Y")] = "Yes"
head(merged_file)
```

## Save the merged data

```{r,eval=FALSE}
# write.xlsx(x=merged_file,file=paste0(dataDir,"PatientsData_merged_",today,".xlsx"))
```

## Descriptive table

* samples with no missing values (n=685)    
* zip code with at least 5 samples (n=883)

```{r}
sample0 = which(apply(merged_file,1,FUN=function(x){sum(is.na(x))})==0) # samples with no missing values
cat(paste("Number of samples with no missing =",length(sample0)),"\n")

merged_file0 = merged_file[sample0,]
dim(merged_file0)
table(merged_file0$zip)
zipcode0 = sort(unique(merged_file0$zip))
# zipcode0 = sort(names(table(merged_file0$zip)[which(table(merged_file0$zip)>5)]))
# merged_file0 = merged_file0[which(merged_file0$zip %in% zipcode),]
# table(merged_file0$zip)

```

* samples with no missing values in POVERTY (n=1410)

```{r}
sample1 = which(apply(merged_file[,c(3,4,31:34)],1,FUN=function(x){sum(is.na(x))})==0) # samples with no missing values in POVERTY
cat(paste("Number of samples with no missing =",length(sample1)),"\n")

merged_file1 = merged_file[sample1,]
dim(merged_file1)
table(merged_file1$zip)
zipcode1 = sort(unique(merged_file1$zip))
# zipcode1 = sort(names(table(merged_file1$zip)[which(table(merged_file1$zip)>5)]))
# merged_file1 = merged_file1[which(merged_file1$zip %in% zipcode1),]
# table(merged_file1$zip)
```

# PCM data (pain perception)

Goal. PCM + dx_date_use + first_chemo + first_surgery

* Remove duplicates from PCM file (N=3835)

```{r}
rm(pcm_file)
pcm_file_original = read.xlsx(xlsxFile=paste0(dataDir,"pcm.xlsx"),sheet=1,detectDates=TRUE)
pcm_file = pcm_file_original
dim(pcm_file)
length(unique(pcm_file$profile_key))

pcm_uniq1 = distinct(pcm_file)
dim(pcm_uniq1)
length(unique(pcm_uniq1$profile_key))
```

* Remove samples that are not in PatientVar profile

```{r,echo=FALSE}
mergedData=read.xlsx(xlsxFile=paste0(dataDir,"PatientsData_merged_2020-07-15.xlsx"),sheet=1,detectDates=TRUE)

cat(paste("Number of patients in PCM profile but not in PatientVar profile = ",length(which(! unique(pcm_uniq1$profile_key) %in% mergedData$profile_key))),"\n")
cat(paste("Number of patients in PatientVar profile but not in PCM profile = ",length(which(! mergedData$profile_key %in% unique(pcm_uniq1$profile_key)))),"\n")

pcm_uniq2 = pcm_uniq1[which(pcm_uniq1$profile_key %in% mergedData$profile_key),]
dim(pcm_uniq2)
length(unique(pcm_uniq2$profile_key))
```

* Remove NAs from PCM file

```{r,echo=FALSE}
length(which(is.na(pcm_uniq2$a_CCMA06)))
pcm_uniq3 = pcm_uniq2[which(! is.na(pcm_uniq2$a_CCMA06)),]
dim(pcm_uniq3)
length(unique(pcm_uniq3$profile_key))

cat(paste("Number of patients in PatientVar profile but not in PCM profile = ",length(which(! mergedData$profile_key %in% unique(pcm_uniq3$profile_key)))),"\n")

## Final
pcm_file = pcm_uniq3
profile_key = mergedData$profile_key
```

* Put dx_date_use to the closest survey date in PCM file.

```{r,echo=FALSE}
dx_date = as.Date(rep(NA,dim(pcm_file)[1]))
dx_diff_days = rep(NA,dim(pcm_file)[1])

dx_date_na_num = 0
for (i in 1:length(unique(pcm_file$profile_key))) {
  x = unique(pcm_file$profile_key)[i]
  temp_rows=which(pcm_file$profile_key==x)
  # pcm_file[temp_rows,]
  if (! is.na(mergedData$dx_date_use[which(mergedData$profile_key==x)])) {
    diff_days0=pcm_file$SURVEY_DATE[temp_rows]-mergedData$dx_date_use[which(mergedData$profile_key==x)]
    dx_diff_days[temp_rows] = diff_days0
    sel_row = temp_rows[which(abs(diff_days0)==min(abs(diff_days0),na.rm=TRUE))][1]
    dx_date[sel_row] = mergedData$dx_date_use[which(mergedData$profile_key==x)]
  } else {
    dx_date_na_num = dx_date_na_num + 1    
  }
  # cat(paste(i),"\n")
}
pcm_final = cbind(pcm_file,dx_date,dx_diff_days)

cat(paste("Number of patients in PCM with dx_date not available =",dx_date_na_num),"\n")
cat(paste("Number of patients in PCM with dx_date available =",length(which(!is.na(pcm_final$dx_date)))),"\n")
rm(diff_days0)
```

* Put first_chemo to the closest survey date in PCM file.

```{r,echo=FALSE}
first_chemo_date = as.Date(rep(NA,dim(pcm_file)[1]))
chemo_diff_days = rep(NA,dim(pcm_file)[1])

first_chemo_na_num = 0
for (i in 1:length(unique(pcm_file$profile_key))) {
  x = unique(pcm_file$profile_key)[i]
  # x = "112907"
  temp_rows=which(pcm_file$profile_key==x)
  # pcm_file[temp_rows,]
  if (! is.na(mergedData$first_chemo[which(mergedData$profile_key==x)])) {
    diff_days0=pcm_file$SURVEY_DATE[temp_rows]-as.Date(mergedData$first_chemo[which(mergedData$profile_key==x)])
    chemo_diff_days[temp_rows] = diff_days0
    sel_row = temp_rows[which(abs(diff_days0)==min(abs(diff_days0),na.rm=TRUE))][1]
    first_chemo_date[sel_row] = mergedData$first_chemo[which(mergedData$profile_key==x)]
  } else {
    first_chemo_na_num = first_chemo_na_num + 1    
  }
  # cat(paste(i),"\n")
}
pcm_final = cbind(pcm_final,first_chemo_date,chemo_diff_days)

cat(paste("Number of patients in PCM with first_chemo not available =",first_chemo_na_num),"\n")
cat(paste("Number of patients in PCM with first_chemo available =",length(which(!is.na(pcm_final$first_chemo_date)))),"\n")

```

* Put first_surgery to the closest survey date in PCM file.

```{r,echo=FALSE}
first_surgery_date = as.Date(rep(NA,dim(pcm_file)[1]))
surgery_diff_days = rep(NA,dim(pcm_file)[1])

first_surgery_na_num = 0
for (i in 1:length(unique(pcm_file$profile_key))) {
  x = unique(pcm_file$profile_key)[i]
  # x = "112907"
  temp_rows=which(pcm_file$profile_key==x)
  # pcm_file[temp_rows,]
  if (! is.na(mergedData$first_surgery[which(mergedData$profile_key==x)])) {
    diff_days0=pcm_file$SURVEY_DATE[temp_rows]-as.Date(mergedData$first_surgery[which(mergedData$profile_key==x)])
    surgery_diff_days[temp_rows] = diff_days0
    sel_row = temp_rows[which(abs(diff_days0)==min(abs(diff_days0),na.rm=TRUE))][1]
    first_surgery_date[sel_row] = mergedData$first_surgery[which(mergedData$profile_key==x)]
  } else {
    first_surgery_na_num = first_surgery_na_num + 1    
  }
  # cat(paste(i),"\n")
}
pcm_final = cbind(pcm_final,first_surgery_date,surgery_diff_days)

cat(paste("Number of patients in PCM with first_surgery not available =",first_surgery_na_num),"\n")
cat(paste("Number of patients in PCM with first_surgery available =",length(which(!is.na(pcm_final$first_surgery_date)))),"\n")

```

```{r,echo=FALSE}
head(pcm_final)
```

```{r, echo=FALSE}
hist(pcm_final$dx_diff_days[which(!is.na(pcm_final$dx_date))]/30,breaks=200,
     xlab="Months")

hist(pcm_final$dx_diff_days[which(!is.na(pcm_final$dx_date))]/30,breaks=200,
     xlab="Months",xlim=c(-12,12))

cat(paste("Number of patients within 30-day window =",length(which(abs(pcm_final$dx_diff_days[which(!is.na(pcm_final$dx_date))])<30))),"\n")
cat(paste("Number of patients within 60-day window =",length(which(abs(pcm_final$dx_diff_days[which(!is.na(pcm_final$dx_date))])<60))),"\n")
cat(paste("Number of patients within 90-day window =",length(which(abs(pcm_final$dx_diff_days[which(!is.na(pcm_final$dx_date))])<90))),"\n")

```

```{r,eval=FALSE}
# write.xlsx(x=pcm_final,file=paste0(dataDir,"PCM_Dates_merged_",today,".xlsx"))
```

### 2020-07-24 manuscript_code_20200722.Rmd
figure updated

---
title: "Code for Breast Cancer Pain Study"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 12
    fig_height: 6
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE,autodeep=TRUE,cache.comments=FALSE,warning=FALSE,message=FALSE,cache.lazy=FALSE)
```

# Data pre-processing

```{r,echo=FALSE}
library(openxlsx)
library(lattice)
library(quantreg)
library(ggplot2)
library(pscl)
library(boot)
library(MASS)
library(sjPlot)
library(nonnest2)
library(vcd)
library(AER)
library(faraway)
library(tidyverse)
library(table1)
library(RColorBrewer)
# library(ggpubr)

```

```{r,echo=FALSE}
params = as.list(NULL)
params$days.cutoff = 12
params$bin.cutoff = 0
params$race = "All"
params$occupation = FALSE
params$patientsData = "PatientsData_merged_2020-07-15.xlsx"
params$PainData = "PCM_Dates_merged_2020-06-29.xlsx"

days.cutoff = as.numeric(params$days.cutoff)*(365/12)
bin.cutoff = as.numeric(params$bin.cutoff) # from what number, pain considered as "no pain"?
export.data = FALSE # save merged data as excel format?

## Load data
patients_file=read.xlsx(xlsxFile=as.character(params$patientsData),sheet=1,detectDates=TRUE)
pcm_file=read.xlsx(xlsxFile=as.character(params$PainData),sheet=1,detectDates=TRUE)

colnames(patients_file)[which(colnames(patients_file)=="POVERTY")] = "Poverty"
colnames(patients_file)[which(colnames(patients_file)=="VACANCY")] = "Vacancy"
colnames(patients_file)[which(colnames(patients_file)=="WALKABILITY")] = "Walkability"
colnames(patients_file)[which(colnames(patients_file)=="UNEMPLOYMENT")] = "Unemployment"
colnames(pcm_file)[which(colnames(pcm_file)=="a_CCMA06")] = "Pain"
```

**_Patients selection_**.

* Cleaned datasets with new dx_date (initial diagnosis date) was released - R21_Sample_wDxDate.     
* Only TN considered. Other states were excluded. (3836 $\rightarrow$ 2628)      
* Some duplicated samples were removed.   
    - There are 43 duplicated profile keys in the merged data set.     
    - e.g. profile key==90006  
* There are a lot of missing values in the important variables such as Marital status, Occupation, Primary Payer, POVERTY, and etc. The missing values are systematic rather than just random, for example, if one patient is missing in one variable then it is also missing in the other variables. Therefore, we will only include the patients with no missing values in any of the important covariates.

* missing, NA -> Unknown

**_Pain data (PCM) selection_**.

* Pain variable (a_CCMA06) observed after at least 60 days prior to dx_date only considered.   
* Pain variable (a_CCMA06) within 365 days after the initial diagnosis date (dx_date) has been chosen.     
* Mean was taken for Pain by patients.

**_Poverty variable collection_**.

* "joined_zipcode" file merged to get data about poverty, vacancy, unemployment levels for more zip codes.   
* Poverty data (from "joined_zipcode")

    - For poverty variable, group mean used for each zip code.   
    - Only 43 zip codes are documented in poverty file. On the other hand, there are 165 zip codes (TN) in Patients file. This produced many missing values for the poverty variable.

**_Further data selection_**.

* Race = "Other" or "Unknown" removed. (171 Samples excluded)
* NA's in Pain variable removed. (247 samples excluded)

```{r,echo=FALSE}
pcm_profile_key=sort(unique(pcm_file$profile_key))
patients_profile_key=sort(unique(patients_file$profile_key))

cat(paste("Number of patients in the cohort (PatientsData) =",length(patients_profile_key)),"\n")
pcm_subset = pcm_file[which((abs(pcm_file$dx_diff_days)<days.cutoff) & (pcm_file$dx_diff_days>(-60))),]

pcm_new = cbind(aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(mean(x))}),
                pcm_Q1st = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(quantile(x)[2])})[,2],
                pcm_Q2nd = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(quantile(x)[3])})[,2],
                pcm_Q3rd = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){round(quantile(x)[4])})[,2],
                pcm_min = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){min(x)})[,2],
                pcm_max = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){max(x)})[,2],
                pcm_bin = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){as.numeric((min(x)>bin.cutoff))})[,2])
colnames(pcm_new) = c(colnames(pcm_new)[1],paste(c("mean","Q1st","Q2nd","Q3rd","min","max","bin"),colnames(pcm_new)[2],sep="."))

pcm_count = aggregate(. ~ profile_key, pcm_subset[,c(1,3)], FUN=function(x){length(x)})
colnames(pcm_count)[2] = "count.Pain"
pcm_var = merge(x=pcm_new,y=pcm_count,by="profile_key",all=T)
profile_key = pcm_var$profile_key

## Add chemo variable: afchemo / bfchemo (after/before chemo)
afchemo = bfchemo = matrix(NA,ncol=ncol(pcm_var),nrow=length(profile_key))
afchemo[,1] = bfchemo[,1] = profile_key
colnames(bfchemo) = paste("bfchemo",colnames(pcm_var),sep=".")
colnames(afchemo) = paste("afchemo",colnames(pcm_var),sep=".")
colnames(bfchemo)[1] = colnames(pcm_var)[1]
colnames(afchemo)[1] = colnames(pcm_var)[1]

for (i in 1:length(profile_key)) {
  x = profile_key[i]
  temp_rows=which(pcm_subset$profile_key==x)
  pcm_subset[temp_rows,]

  diff.date = (pcm_subset$dx_diff_days[temp_rows] - pcm_subset$chemo_diff_days[temp_rows])[1] # chemo.date - dx.date
  if (is.na(diff.date)) {
    next
  } else if ((diff.date>(-365)) & (diff.date<(days.cutoff))) {
    bfchemo.table = pcm_subset[temp_rows[which(pcm_subset$chemo_diff_days[temp_rows]<=0)],]
    afchemo.table = pcm_subset[temp_rows[which(pcm_subset$chemo_diff_days[temp_rows]>0)],]

    if (dim(bfchemo.table)[1]>0) {
      bfchemo[i,2:ncol(bfchemo)] = c(round(mean(bfchemo.table$Pain)),
                                     round(quantile(bfchemo.table$Pain)[2]),
                                     round(quantile(bfchemo.table$Pain)[3]),
                                     round(quantile(bfchemo.table$Pain)[4]),
                                     min(bfchemo.table$Pain),
                                     max(bfchemo.table$Pain),
                                     as.numeric((min(bfchemo.table$Pain)>bin.cutoff)),
                                     dim(bfchemo.table)[1])
    }
    if (dim(afchemo.table)[1]>0) {
      afchemo[i,2:ncol(afchemo)] = c(round(mean(afchemo.table$Pain)),
                                     round(quantile(afchemo.table$Pain)[2]),
                                     round(quantile(afchemo.table$Pain)[3]),
                                     round(quantile(afchemo.table$Pain)[4]),
                                     min(afchemo.table$Pain),
                                     max(afchemo.table$Pain),
                                     as.numeric((min(afchemo.table$Pain)>bin.cutoff)),
                                     dim(afchemo.table)[1])
    }
  }
  # cat(paste(i),"\n")
}
pcm_var2 = merge(x=pcm_var,y=bfchemo,by="profile_key",all=T)
pcm_var2 = merge(x=pcm_var2,y=afchemo,by="profile_key",all=T)
# head(pcm_var2,n=20)

## Add surgery variable: afsurgery / bfsurgery (after/before surgery)
afsurgery = bfsurgery = matrix(NA,ncol=ncol(pcm_var),nrow=length(profile_key))
afsurgery[,1] = bfsurgery[,1] = profile_key
colnames(bfsurgery) = paste("bfsurgery",colnames(pcm_var),sep=".")
colnames(afsurgery) = paste("afsurgery",colnames(pcm_var),sep=".")
colnames(bfsurgery)[1] = colnames(pcm_var)[1]
colnames(afsurgery)[1] = colnames(pcm_var)[1]

for (i in 1:length(profile_key)) {
  x = profile_key[i]
  temp_rows=which(pcm_subset$profile_key==x)
  pcm_subset[temp_rows,]

  diff.date = (pcm_subset$dx_diff_days[temp_rows] - pcm_subset$surgery_diff_days[temp_rows])[1] # surgery.date - dx.date
  if (is.na(diff.date)) {
    next
  } else if ((diff.date>(-365)) & (diff.date<(days.cutoff))) {
    bfsurgery.table = pcm_subset[temp_rows[which(pcm_subset$surgery_diff_days[temp_rows]<=0)],]
    afsurgery.table = pcm_subset[temp_rows[which(pcm_subset$surgery_diff_days[temp_rows]>0)],]

    if (dim(bfsurgery.table)[1]>0) {
      bfsurgery[i,2:ncol(bfsurgery)] = c(round(mean(bfsurgery.table$Pain)),
                                     round(quantile(bfsurgery.table$Pain)[2]),
                                     round(quantile(bfsurgery.table$Pain)[3]),
                                     round(quantile(bfsurgery.table$Pain)[4]),
                                     min(bfsurgery.table$Pain),
                                     max(bfsurgery.table$Pain),
                                     as.numeric((min(bfsurgery.table$Pain)>bin.cutoff)),
                                     dim(bfsurgery.table)[1])
    }
    if (dim(afsurgery.table)[1]>0) {
      afsurgery[i,2:ncol(afsurgery)] = c(round(mean(afsurgery.table$Pain)),
                                     round(quantile(afsurgery.table$Pain)[2]),
                                     round(quantile(afsurgery.table$Pain)[3]),
                                     round(quantile(afsurgery.table$Pain)[4]),
                                     min(afsurgery.table$Pain),
                                     max(afsurgery.table$Pain),
                                     as.numeric((min(afsurgery.table$Pain)>bin.cutoff)),
                                     dim(afsurgery.table)[1])
    }
  }
  # cat(paste(i),"\n")
}

pcm_var2 = merge(x=pcm_var2,y=bfsurgery,by="profile_key",all=T)
pcm_var2 = merge(x=pcm_var2,y=afsurgery,by="profile_key",all=T)
# head(pcm_var2,n=20)

dmd = merge(pcm_var2,patients_file,by="profile_key",all=T)
# if (export.data) {
#   write.xlsx(x=dmd,file=paste0(dataDir,"PCM_PatientsVar_merged_DX",days.cutoff,"daysRule_",today,".xlsx"))
# }

```

```{r,echo=FALSE}
## Define factor variables:
# table(dmd$Tobacco.History)
dmd$Tobacco.History = factor(dmd$Tobacco.History,levels=c("Never","Smoked","Unknown"))

# table(dmd$Histology)
dmd$Histology = factor(dmd$Histology,levels=c("ductal","lobular","Other","Unknown"))

# table(dmd$Recurrence)
dmd$Recurrence = factor(dmd$Recurrence,levels=c("disease-free","Recurrence","Unknown"))

# table(dmd$Primary.Site.1)
dmd$Primary.Site.1 = factor(dmd$Primary.Site.1,levels=c("Lower","Central","Upper","Unknown"))

# table(dmd$Primary.Site.2)
dmd$Primary.Site.2 = factor(dmd$Primary.Site.2,levels=c("Inner","Midline","Outer","Overlapping","Unknown"))

# table(dmd$HER2)
dmd$HER2[(dmd$HER2 %in% c("Borderline","Test.not.done","Unknown"))] = "Unknown"
dmd$HER2 = factor(dmd$HER2,levels=c("Negative.Normal","Positive.Elevated","Unknown"))

# table(dmd$ER)
dmd$ER[(dmd$ER %in% c("Test.not.done","Unknown"))] = "Unknown"
dmd$ER = factor(dmd$ER,levels=c("Negative.Normal","Positive.Elevated","Unknown"))

# table(dmd$PR)
dmd$PR[(dmd$PR %in% c("Test.not.done","Unknown"))] = "Unknown"
dmd$PR = factor(dmd$PR,levels=c("Negative.Normal","Positive.Elevated","Unknown"))

# table(dmd$Metastasis)
dmd$Metastasis = factor(dmd$Metastasis,levels=c("No","Yes"))

# Relevel factor variables
dmd$race.f = relevel(factor(dmd$race.f),"White")
# table(dmd0$race.f)

dmd$marital.status.f = relevel(factor(dmd$marital.status.f),"Married")
dmd$marital.status.f = factor(dmd$marital.status.f, levels=c("Married","Single","Divorced","Widowed","Unknown"))
# table(dmd$marital.status.f)

dmd$Usual.Occupation.f = relevel(factor(dmd$Usual.Occupation.f),"Employed")
# table(dmd0$Usual.Occupation.f)

dmd$Primary.Payer.f = relevel(factor(dmd$Primary.Payer.f),"Insurance")
# table(dmd0$Primary.Payer.f)

# table(dmd$race.f)

## Usual Occupation
occupation = dmd$Usual.Occupation.f
occupation[which(occupation=="Disabled")] = "Unemployed"
if (params$occupation) {
  occupation[which(occupation=="Retired")] = "Unemployed"
}
occupation.f = factor(occupation)
dmd$Usual.Occupation.f2 = occupation.f
# table(dmd$Usual.Occupation.f2)

## Primary Payer
insurance = dmd$Primary.Payer.f
insurance[which(insurance=="Private")] = "Insurance"
insurance.f = factor(insurance)
dmd$Primary.Payer.f2 = insurance.f
# table(dmd$Primary.Payer.f2)

## Marital Status
marriage = dmd$marital.status.f
marriage[which(marriage=="Widowed")] = "Single"
marriage.f = factor(marriage)
dmd$marital.status.f2 = marriage.f
# table(dmd$marital.status.f2)

## Poverty -> factor (low/high)
Poverty = rep("low",dim(dmd)[1])
Poverty[which(dmd$Poverty>median(dmd$Poverty,na.rm=T))] = "high"
Poverty.f = factor(Poverty)
Poverty.f = relevel(Poverty.f,"low")
dmd$Poverty.f = Poverty.f
# table(dmd$Poverty.f)

## Stage
dmd$stage.f = factor(dmd$stage.f)
dmd$stage.f = relevel(dmd$stage.f,"1")

## Age
age = rep("Low",dim(dmd)[1])
age[which(dmd$age.f==1)] = "High"
dmd$age.f = factor(age)
dmd$age.f = relevel(dmd$age.f,"Low")

## Chemo -> factor (0=no, 1=yes)
chemo = rep("Untreated",dim(dmd)[1])
chemo[which(dmd$afchemo.count.Pain>0)] = "Treated"
chemo.f = factor(chemo)
chemo.f = relevel(chemo.f,"Treated")
dmd$chemo.f = chemo.f
# table(dmd$chemo.f)

## surgery -> factor (0=no, 1=yes)
surgery = rep("Untreated",dim(dmd)[1])
surgery[which(dmd$afsurgery.count.Pain>0)] = "Treated"
surgery.f = factor(surgery)
surgery.f = relevel(surgery.f,"Treated")
dmd$surgery.f = surgery.f
# table(dmd$surgery.f)

if (! params$race=="All") {
  dmd = dmd[which(dmd$race.f==params$race),]
}

unknownRaceN = length(which(dmd$race.f %in% c("Other","Unknown")))
dmd = dmd[-which(dmd$race.f %in% c("Other","Unknown")),]
cat(paste("Unknown and Other in Race removed (",unknownRaceN,"), Black and White remained N =",dim(dmd)[1]),"\n")
dmd$race.f = factor(dmd$race.f)

## Re-name variables
# colnames(cdmd.pvt)[which(grepl(pattern="surgery",colnames(cdmd.pvt))==T)]
colnames(dmd)[which(colnames(dmd)=="Usual.Occupation.f2")] = "Occupation"
colnames(dmd)[which(colnames(dmd)=="marital.status.f2")] = "Marital.status"
colnames(dmd)[which(colnames(dmd)=="Primary.Payer.f2")] = "Insurance.type"
colnames(dmd)[which(colnames(dmd)=="race.f")] = "Race"
colnames(dmd)[which(colnames(dmd)=="stage.f")] = "Cancer.stage"
colnames(dmd)[which(colnames(dmd)=="age.f")] = "Age"
colnames(dmd)[which(colnames(dmd)=="chemo.f")] = "Chemo"
colnames(dmd)[which(colnames(dmd)=="surgery.f")] = "Surgery"
colnames(dmd)[which(colnames(dmd)=="Tobacco.History")] = "Tobacco.history"
```

```{r,echo=FALSE}
## Select samples with Poverty values
dmd.pvt = dmd[which(!is.na(dmd$Poverty)==T),]
zipcode.pvt = sort(unique(dmd.pvt$zip))
cat(paste("Number of patients with Poverty reported =",dim(dmd.pvt)[1]),"\n")

dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
cat(paste("Three outliers excluded"),"\n")

cdmd.pvt = dmd.pvt
colnames(cdmd.pvt)[which(grepl(pattern="mean",colnames(cdmd.pvt))==T)] = c("Pain","Pain.bfchemo","Pain.afchemo","Pain.bfsurgery","Pain.afsurgery")
NANs = length(which(is.na(cdmd.pvt$Pain)))
cdmd.pvt = cdmd.pvt[which(!is.na(cdmd.pvt$Pain)),]
cat(paste("NAs in Pain removed (final dataset) =",dim(cdmd.pvt)[1]," (NAs =",NANs,")"),"\n")
cat(paste("Number of unique zip code =",length(unique(cdmd.pvt$zip))),"\n")

cdmd.pvt$Poverty.q = rep("Low",nrow(cdmd.pvt))
cdmd.pvt$Poverty.q[which(cdmd.pvt$Poverty>quantile(cdmd.pvt$Poverty,p=0.3))] = "Medium"
cdmd.pvt$Poverty.q[which(cdmd.pvt$Poverty>quantile(cdmd.pvt$Poverty,p=0.7))] = "High"
```

```{r}
final.data <- 
  cdmd.pvt %>%
  select("Pain",
         "Poverty","Vacancy",
         "Occupation","Marital.status","Insurance.type",
         "Cancer.stage","Chemo","Surgery",
         "Race","Age",
         "Histology",
         "Primary.Site.1","Primary.Site.2",
         "HER2","ER","PR","Metastasis")
```

# Summary of variables

## Table 1

```{r,echo=FALSE,eval=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")

table1(as.formula(paste(" ~ ",paste(c(respvar,explvars),collapse=" + "))), data=final.data)
```

```{r,echo=FALSE}
respvar = "Pain"
explvars = c("Pain","Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

table1(as.formula(paste(" ~ ",paste(c(respvar,explvars),collapse=" + "))), data=final.data)
```

```{r,echo=FALSE,eval=FALSE}
table(final.data)
summary(final.data)
do.call("rbind",apply(final.data,2,get_summary))

get_summary = function(x) {
  tab = table(x)
  names(tab)

  summ = t(t(table(x)))
  names(summ)
}
x = final.data$Marital.status
t(t(table(final.data$Marital.status)))

## Try htmlTable
htmlTable(table.count,
          rgroup = c("","",gsub(pattern="[.]",replacement=" ",x=explvars[2:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,4,4,3,3,2,2,2,2,3),
          cgroup = c("Unadjusted","Adjusted"),
          n.cgroup = c(3,3))

```

## Distributions

**_Pain variable (Pain)_**:

* For each patient, multiple pains were reported at different dates after their initial diagnosis dates. As a representative value of these multiple observations, average was computed for each patient within the 365-day window.

```{r,echo=FALSE,fig.width=8}
# pdf(6,6,file=paste("Pain_hist.pdf"))
barplot(table(cdmd.pvt$Pain),main=paste("Pain intensity"),ylab="Frequency")
# legend("topright",bty="n",legend=c(paste("Mean   =",round(mean(cdmd.pvt$Pain),digits=2)),
                                   # paste("Median =",round(median(cdmd.pvt$Pain),digits=2))))
# dev.off()

```

**_Poverty variable_**:

* Poverty was measured at zip code level. Since we have 34 different zip code, Poverty variable also has 34 unique values each of which corresponds to each zip code.

```{r,echo=FALSE,fig.width=8}

hist(cdmd.pvt$Poverty,main="Poverty")
cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
cat(paste("Median of Poverty =",round(median(cdmd.pvt$Poverty),digits=3)),"\n")

```

**_Vacancy variable_**:

```{r,echo=FALSE,fig.width=8}

hist(cdmd.pvt$Vacancy,main="Vacancy")
cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
cat(paste("Median of Vacancy =",round(median(cdmd.pvt$Vacancy),digits=3)),"\n")

```

## Poverty

* Pain ~ Poverty

```{r,echo=FALSE}
p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Poverty = aggregate(Poverty ~ zip,cdmd.pvt,mean)$Poverty,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Poverty ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Poverty,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Poverty))

# par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))

pdf(6,6,file=paste("Pain_poverty_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Poverty,y=meandata$Pain,
     xlim=c(0,70),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Poverty", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

# points(x=meandata$Poverty,y=meandata$Pain,
#        cex=cex,lwd=1.5)
x=seq(0,70,by=1)
points(x=x,y=fhat(x),type="l",lwd=3,
       col=brewer.pal(n=9,"YlOrRd")[7])
dev.off()

# par(mfrow=c(3,3))
pdf(6,6,file=paste("Pain_poverty_legend.pdf"))
par(mar=c(4,4,2,2))
plot(x=c(0,1),y=c(0,1))
cbind(meandata$size,cex)

x = c(1,20,40,60,80)
y = c(1,1.5,2.5,4,6)
points(x=rep(0.4,length(x)),
       y=seq(0.8,0.4,length.out=length(x)),
       cex=y,lwd=2.5,col=brewer.pal(n=11,"BrBG")[9])
text(x=rep(0.55,length(x)),
     y=seq(0.8,0.4,length.out=length(x)),
     labels=x,cex=1.5)
dev.off()

```

## Vacancy

* Pain ~ Vacancy

```{r,echo=FALSE}
p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
# summary(p.glm)
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))

# par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))

pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
par(mar=c(4,4,2,2))
plot(x=meandata$Vacancy,y=meandata$Pain,
     xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
     axes=F,
     xlab=NA,ylab=NA,
     cex=cex,lwd=2.5)
box()
abline(0, 0, lty=1, col="grey")
abline(v=0, lty=1, col="grey")
# axis(side=1, tck=-.015,labels=NA)
# axis(side=1, lwd=0, line=-1, cex=0.2,cex.axis=0.9)
axis(side=1, tck=-0.015,lwd=0,line=-.5,cex.axis=1.5)
axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
mtext(side=1, "Vacancy", line=2.5, cex=1.5)
mtext(side=2, "Pain average", line=2, cex=1.5)

x=seq(0,53,by=1)
points(x=x,y=fhat(x),type="l",lwd=3,
       col=brewer.pal(n=9,"YlOrRd")[7])
dev.off()

```

```{r,echo=FALSE}
p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
# summary(p.glm)
# summary(zeroinfl(Pain ~ Poverty, data = cdmd.pvt, dist="poisson"))
fhat = function(x) {
  exp(p.glm$coefficients[1]+p.glm$coefficients[2]*x)
}

meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
                      Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
# hist(round(meandata$Pain))
# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))

par(mfrow=c(1,2))
cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
# pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
plot(x=meandata$Vacancy,y=meandata$Pain,cex=cex,xlab="Vacancy",ylab="Pain average")
# plot(x=aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
#      y=aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
#      xlab="Vacancy",ylab=paste(pain.type,"Pain mean within Zip"))
x=seq(0,80,by=1)
points(x=x,y=fhat(x),type="l",lwd=2)
# dev.off()
```

```{r,echo=FALSE}
respvar = "Pain"

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age")
explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age")
full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

```{r,echo=FALSE,eval=FALSE}

## Get the ordered explanatory variables in full model
full.model = zeroinfl(full.formula, data = final.data, dist="poisson")
full.summary = round(cbind(exp(coef(full.model)),exp(confint(full.model)),do.call("rbind",coef(summary(full.model)))[,c(4)]),digits=3)
table.explvars = rownames(full.summary)

final.1 = zeroinfl(full.formula.1, data = final.data, dist="poisson")
final.2 = zeroinfl(full.formula.2, data = final.data, dist="poisson")

# tab_model(final, digits = 3,
#           title = "Final",
#           show.zeroinf=TRUE,show.intercept=TRUE,
#           show.se=TRUE, show.ci=FALSE,collapse.se=FALSE, 
#           show.p = TRUE, p.style="numeric_stars")

```

```{r,echo=FALSE}
# exp(confint(final)) # confidence interval
final.adj.1 = round(cbind(exp(coef(final.1)),exp(confint(final.1)),do.call("rbind",coef(summary(final.1)))[,c(4)]),digits=3)
colnames(final.adj.1)[c(1,4)] = c("Estimate","P")
final.adj.2 = round(cbind(exp(coef(final.2)),exp(confint(final.2)),do.call("rbind",coef(summary(final.2)))[,c(4)]),digits=3)
colnames(final.adj.2)[c(1,4)] = c("Estimate","P")

## Final model (Unadjusted)
final.unadj.list = as.list(NULL)
formula.base = Pain ~ Poverty
for (i in 1:length(explvars)) {
  formula.new = as.formula(paste(sub("Poverty",replacement=explvars[i],as.character(formula.base))[c(2,3)],collapse=" ~ "))
  model.unadj = zeroinfl(formula.new, data = final.data, dist="poisson")
  summary.unadj = round(cbind(exp(coef(model.unadj)),exp(confint(model.unadj)),do.call("rbind",coef(summary(model.unadj)))[,c(4)]),digits=3)
  colnames(summary.unadj)[c(1,4)] = c("Estimate","P")
  colnames(summary.unadj) = paste("Unadj",colnames(summary.unadj),sep=".")
  final.unadj.list[[i]] <- summary.unadj
  rm(summary.unadj)
}
final.unadj = do.call("rbind",final.unadj.list)
final.unadj = final.unadj[which(grepl(pattern="Intercept",rownames(final.unadj))==F),]

## Merge
final.adj.1 = as.data.frame(final.adj.1)
final.adj.1$vars = rownames(final.adj.1)
final.adj.2 = as.data.frame(final.adj.2)
final.adj.2$vars = rownames(final.adj.2)
final.unadj = as.data.frame(final.unadj)
final.unadj$vars = rownames(final.unadj)

final.table0 = merge(final.adj.1,final.adj.2,by="vars",all=T)
final.table1 = merge(final.table0,final.unadj,by="vars",all=T)
# which(! table.explvars %in% final.table0$vars)
# which(! final.table0$vars %in% table.explvars)
final.table = final.table1[match(table.explvars,final.table1$vars),c(10:13,2:9)]
rownames(final.table) = table.explvars

final.table$Unadj.CI = paste(final.table$`Unadj.2.5 %`,final.table$`Unadj.97.5 %`,sep="-")
final.table$CI.x = paste(final.table$`2.5 %.x`,final.table$`97.5 %.x`,sep="-")
final.table$CI.y = paste(final.table$`2.5 %.y`,final.table$`97.5 %.y`,sep="-")
final.table <- final.table %>%
  select(Unadj.Estimate,Unadj.CI,Unadj.P,Estimate.x,CI.x,P.x,Estimate.y,CI.y,P.y)

transform.p = function(x) {
  y = as.character(x)
  y[which(y==0)] = "<.001"
  return(y)
}
final.table$Unadj.P = transform.p(final.table$Unadj.P)
final.table$P.x = transform.p(final.table$P.x)
final.table$P.y = transform.p(final.table$P.y)
```

## Poisson count model
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table

table.count = final.table[which(grepl("count",rownames(final.table))),]
rownames(table.count) = sapply(rownames(table.count),function(x) unlist(strsplit(x,split="count_"))[2]) # extract count model

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.count,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
# library(kableExtra)
# kable(table.count) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))

```

## Inflation model
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))

# kable(table.zero) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))
```

# Basic model + more biomarker variables

```{r,echo=FALSE}
respvar = "Pain"

explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR","Metastasis")

explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
               "Cancer.stage","Chemo","Surgery",
               "Race","Age",
               "Histology",
               "Primary.Site.1","Primary.Site.2",
               "HER2","ER","PR","Metastasis")

explvars = c("Poverty","Vacancy",
             "Occupation","Marital.status","Insurance.type",
             "Cancer.stage","Chemo","Surgery",
             "Race","Age",
             "Histology",
             "Primary.Site.1","Primary.Site.2",
             "HER2","ER","PR","Metastasis")

full.formula.1 = as.formula(paste(respvar," ~ ",paste(explvars.1,collapse=" + ")))
full.formula.2 = as.formula(paste(respvar," ~ ",paste(explvars.2,collapse=" + ")))
full.formula = as.formula(paste(respvar," ~ ",paste(explvars,collapse=" + ")))
```

```{r,echo=FALSE}
## Get the ordered explanatory variables in full model
full.model = zeroinfl(full.formula, data = final.data, dist="poisson")
full.summary = round(cbind(exp(coef(full.model)),exp(confint(full.model)),do.call("rbind",coef(summary(full.model)))[,c(4)]),digits=3)
table.explvars = rownames(full.summary)

final.1 = zeroinfl(full.formula.1, data = final.data, dist="poisson")
final.2 = zeroinfl(full.formula.2, data = final.data, dist="poisson")

# tab_model(final, digits = 3,
#           title = "Final",
#           show.zeroinf=TRUE,show.intercept=TRUE,
#           show.se=TRUE, show.ci=FALSE,collapse.se=FALSE, 
#           show.p = TRUE, p.style="numeric_stars")

```

```{r,echo=FALSE}
# exp(confint(final)) # confidence interval
final.adj.1 = round(cbind(exp(coef(final.1)),exp(confint(final.1)),do.call("rbind",coef(summary(final.1)))[,c(4)]),digits=3)
colnames(final.adj.1)[c(1,4)] = c("Estimate","P")
final.adj.2 = round(cbind(exp(coef(final.2)),exp(confint(final.2)),do.call("rbind",coef(summary(final.2)))[,c(4)]),digits=3)
colnames(final.adj.2)[c(1,4)] = c("Estimate","P")

## Final model (Unadjusted)
final.unadj.list = as.list(NULL)
formula.base = Pain ~ Poverty
for (i in 1:length(explvars)) {
  formula.new = as.formula(paste(sub("Poverty",replacement=explvars[i],as.character(formula.base))[c(2,3)],collapse=" ~ "))
  model.unadj = zeroinfl(formula.new, data = final.data, dist="poisson")
  summary.unadj = round(cbind(exp(coef(model.unadj)),exp(confint(model.unadj)),do.call("rbind",coef(summary(model.unadj)))[,c(4)]),digits=3)
  colnames(summary.unadj)[c(1,4)] = c("Estimate","P")
  colnames(summary.unadj) = paste("Unadj",colnames(summary.unadj),sep=".")
  final.unadj.list[[i]] <- summary.unadj
  rm(summary.unadj)
}
final.unadj = do.call("rbind",final.unadj.list)
final.unadj = final.unadj[which(grepl(pattern="Intercept",rownames(final.unadj))==F),]

## Merge
final.adj.1 = as.data.frame(final.adj.1)
final.adj.1$vars = rownames(final.adj.1)
final.adj.2 = as.data.frame(final.adj.2)
final.adj.2$vars = rownames(final.adj.2)
final.unadj = as.data.frame(final.unadj)
final.unadj$vars = rownames(final.unadj)

final.table0 = merge(final.adj.1,final.adj.2,by="vars",all=T)
final.table1 = merge(final.table0,final.unadj,by="vars",all=T)
# which(! table.explvars %in% final.table0$vars)
# which(! final.table0$vars %in% table.explvars)
final.table = final.table1[match(table.explvars,final.table1$vars),c(10:13,2:9)]
rownames(final.table) = table.explvars

final.table$Unadj.CI = paste(final.table$`Unadj.2.5 %`,final.table$`Unadj.97.5 %`,sep="-")
final.table$CI.x = paste(final.table$`2.5 %.x`,final.table$`97.5 %.x`,sep="-")
final.table$CI.y = paste(final.table$`2.5 %.y`,final.table$`97.5 %.y`,sep="-")
final.table <- final.table %>%
  select(Unadj.Estimate,Unadj.CI,Unadj.P,Estimate.x,CI.x,P.x,Estimate.y,CI.y,P.y)

transform.p = function(x) {
  y = as.character(x)
  y[which(y==0)] = "<.001"
  return(y)
}
final.table$Unadj.P = transform.p(final.table$Unadj.P)
final.table$P.x = transform.p(final.table$P.x)
final.table$P.y = transform.p(final.table$P.y)
```

## Poisson count model:
```{r,echo=FALSE}
## tables for publication
library(htmlTable)
# library(Gmisc)
vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table

table.count = final.table[which(grepl("count",rownames(final.table))),]
rownames(table.count) = sapply(rownames(table.count),function(x) unlist(strsplit(x,split="count_"))[2]) # extract count model

table.count <-
  table.count %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.count)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.count)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.count)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.count)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.count)))) %>%
  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.count)))) %>%
  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.count)))) %>%
  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.count)))) %>%
  add_row(,.before=14+which(grepl(pattern="Metastasis",rownames(table.count))))
colnames(table.count) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.count,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))
# library(kableExtra)
# kable(table.count) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))

```

## Inflation model:
```{r,echo=FALSE}
table.zero = final.table[which(grepl("zero",rownames(final.table))),]
rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model

table.zero <-
  table.zero %>%
  add_row(,.before=which(grepl(pattern="Retired",rownames(table.zero)))) %>%
  add_row(,.before=1+which(grepl(pattern="Single",rownames(table.zero)))) %>%
  add_row(,.before=2+which(grepl(pattern="Medicaid",rownames(table.zero)))) %>%
  add_row(,.before=3+which(grepl(pattern="stage2",rownames(table.zero)))) %>%
  add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.zero)))) %>%
  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.zero)))) %>%
  add_row(,.before=14+which(grepl(pattern="Metastasis",rownames(table.zero))))
colnames(table.zero) = c("Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value",
                          "Estimate","95% CI","P-value")

htmlTable(table.zero,
          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
          rnames = tablevars,
          n.rgroup = c(1,1,1,sapply(vars.list,length)),
          cgroup = c("Unadjusted","Adjusted","Adjusted"),
          n.cgroup = c(3,3,3))

# kable(table.zero) %>%
#   kable_styling("striped") %>%
#   add_header_above(c(" " = 1, "Unadjusted" = 3, "Adjusted" = 3))
```

### 2020-07-24 manuscript_code_20200722.Rmd
tables

@@
 
-```{r,echo=FALSE,eval=FALSE}
-table(final.data)
-summary(final.data)
-do.call("rbind",apply(final.data,2,get_summary))
-
-get_summary = function(x) {
-  tab = table(x)
-  names(tab)
-
-  summ = t(t(table(x)))
-  names(summ)
-}
-x = final.data$Marital.status
-t(t(table(final.data$Marital.status)))
-
-## Try htmlTable
-htmlTable(table.count,
-          rgroup = c("","",gsub(pattern="[.]",replacement=" ",x=explvars[2:length(explvars)])),
-          rnames = tablevars,
-          n.rgroup = c(1,1,4,4,3,3,2,2,2,2,3),
-          cgroup = c("Unadjusted","Adjusted"),
-          n.cgroup = c(3,3))
-
-```
-
-## Distributions
+## Figure 2
 
@@
 
-## Poverty
+### Poverty
 
@@
 par(mar=c(4,4,2,2))
-plot(x=c(0,1),y=c(0,1))
-cbind(meandata$size,cex)
-
-x = c(1,20,40,60,80)
-y = c(1,1.5,2.5,4,6)
+plot(x=c(0,1),y=c(0,1),axes=F,cex=0,xlab=NA,ylab=NA)
+# cbind(meandata$size,cex)
+
+x = c(20,40,60,80)
+y = c(1.5,2.5,4,6)
 points(x=rep(0.4,length(x)),
@@
      labels=x,cex=1.5)
+text(x=0.5,y=0.95,
+     labels="Number of patients",cex=1.5)
 dev.off()
@@
 
-## Vacancy
+### Vacancy
 
@@
 ```
+
+# Analysis
+
+## Figure 2
+
+# Significance
+
+```{r,echo=FALSE}
+xlist = c("Poverty.f","marital.status.f","Primary.Payer.f2","Usual.Occupation.f2","race.f","stage.f","age.f","chemo.f","surgery.f")
+
+x = xlist[1]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[2]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[3]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[4]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[5]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[6]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[7]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[8]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+x = xlist[9]
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+               color = x, palette="jco")
+p + stat_compare_means()
+
+```
+
+```{r,echo=FALSE,eval=FALSE}
+## Figure code - boxplot
+
+x = "Poverty"
+tmp = cdmd.pvt[,c("Pain","Poverty.f")]
+colnames(tmp) = c("Pain",x)
+tmp$Poverty = factor(tmp$Poverty)
+tmp$Poverty = relevel(tmp$Poverty,"low")
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Marital.Status"
+tmp <-
+  cdmd.pvt %>%
+  filter(marital.status.f2 %in% c("Married","Single")) %>%
+  select("Pain","marital.status.f2")
+colnames(tmp) = c("Pain",x)
+class(tmp[,2])
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Primary.Payer"
+tmp <-
+  cdmd.pvt %>%
+  filter(Primary.Payer.f2 %in% c("Insurance","Medicaid")) %>%
+  select("Pain","Primary.Payer.f2")
+
+colnames(tmp) = c("Pain",x)
+class(tmp[,2])
+
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Occupation"
+tmp <-
+  cdmd.pvt %>%
+  filter(Usual.Occupation.f2 %in% c("Employed","Retired","Unemployed")) %>%
+  select("Pain","Usual.Occupation.f2")
+
+colnames(tmp) = c("Pain",x)
+class(tmp[,2])
+
+my_comparisons <- list( c("Employed", "Retired"), c("Retired", "Unemployed"), c("Employed", "Unemployed") )
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 15) +
+  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
+
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Race"
+tmp <-
+  cdmd.pvt %>%
+  filter(race.f %in% c("White","Black")) %>%
+  select("Pain","race.f")
+colnames(tmp) = c("Pain",x)
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Age"
+tmp <-
+  cdmd.pvt %>%
+  select("Pain","age.f")
+colnames(tmp) = c("Pain",x)
+tmp$Age[which(tmp$Age==0)] = "low"
+tmp$Age[which(tmp$Age==1)] = "high"
+tmp$Age = factor(tmp$Age)
+tmp$Age = relevel(tmp$Age,"low")
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Surgery"
+tmp <-
+  cdmd.pvt %>%
+  select("Pain","surgery.f")
+colnames(tmp) = c("Pain",x)
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Chemo"
+tmp <-
+  cdmd.pvt %>%
+  select("Pain","chemo.f")
+colnames(tmp) = c("Pain",x)
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 10.5)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = "Stage"
+tmp <-
+  cdmd.pvt %>%
+  select("Pain","stage.f")
+colnames(tmp) = c("Pain",x)
+my_comparisons <- list( c("1", "2"), c("2", "3"), c("1", "3") )
+base <- ggboxplot(tmp, x=x, y="Pain",
+                  color = x, palette="jco",legend="none") +
+  labs(x = NULL, y = "Pain", title = x) + 
+  font("title",size=20,color="black") + 
+  font("x.text", size = 20, color = "black") +
+  font("ylab", size = 20, color = "black") + 
+  font("y.text", size = 20, color = "black") +
+  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+                     label.y = 15) +
+  stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
+p <- base +
+  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+## Table 2
 

### 2020-07-24 manuscript_code_20200722.Rmd
tables

@@
 
-```{r,echo=FALSE}
+```{r,echo=FALSE,eval=FALSE}
 respvar = "Pain"
@@
 
+```{r,echo=FALSE}
+explvars = c("Pain","Poverty","Vacancy",
+             "Occupation","Marital.status","Insurance.type",
+             "Cancer.stage","Chemo","Surgery",
+             "Race","Age",
+             "Histology",
+             "Primary.Site.1","Primary.Site.2",
+             "HER2","ER","PR","Metastasis")
+
+get_summary = function(j) {
+  x = final.data[,j]
+  cname = colnames(final.data)[j]
+  if (class(x)=="numeric") {
+    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
+                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
+    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
+  } else {
+    tmp.sum = t(t(summary(x)))
+    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
+    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
+  }
+  return(tmp.sum)
+}
+table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
+
+get_mean = function(j) {
+  x = final.data[,j]
+  cname = colnames(final.data)[j]
+  if (class(x)=="numeric") {
+    tmp.sum = t(t(c(NA,NA)))
+  } else {
+    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
+  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
+  }
+  return(tmp.sum)
+}
+
+get_pval = function(j) {
+  x = final.data[,j]
+  cname = colnames(final.data)[j]
+  if (class(x)=="numeric") {
+    tmp.sum = t(t(c(1,NA)))
+  } else {
+    tmp.sum = t(t(rep(NA,length(levels(x)))))
+    tmp.sum[1,1] = 1
+  }
+  return(tmp.sum)
+}
+
+table.sum1 = cbind(do.call("rbind",sapply(1:ncol(final.data),get_summary)),
+                   do.call("rbind",sapply(1:ncol(final.data),get_mean)),
+                   do.call("rbind",sapply(1:ncol(final.data),get_pval)))
+colnames(table.sum1) = c("No. (%)","Pain, mean [Q1,Q3]","P-value")
+
+## htmlTable
+library(htmlTable)
+
+rowlabels = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[1])
+htmlTable(table.sum1,
+          rgroup = c(gsub(pattern="[.]",replacement=" ",x=explvars[1:length(explvars)])),
+          rnames = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[2]),
+          n.rgroup = sapply(explvars, function(x) length(which(rowlabels==x))),
+          caption = "Table 1. Baseline characteristics, mean and interquartile range of pain intensity, nonparametric rank test")
+
+```
+
+```{r,echo=FALSE,eval=FALSE}
+library(stats)
+j = 4
+colnames(final.data)[j]
+kruskal.test(final.data[,"Pain"] ~ final.data[,j])
+kruskal.test(final.data[-which(final.data[,j]=="Unknown"),"Pain"] ~ final.data[-which(final.data[,j]=="Unknown"),j])
+å
+kruskal.test(final.data[which(final.data[,j] %in% c("Employed","Unemployed")),"Pain"] ~ final.data[which(final.data[,j] %in% c("Employed","Unemployed")),j])
+
+wilcox.test(final.data[-which(final.data[,j]=="Unknown"),"Pain"] ~ final.data[-which(final.data[,j]=="Unknown"),j])
+
+compare_means(Pain ~ Occupation,data=final.data)
+```
+
 ## Figure 2
@@
 
-```{r,echo=FALSE}
+```{r,echo=FALSE,eval=FALSE}
 p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
@@
 
-```{r,echo=FALSE}
+```{r,echo=FALSE,eval=FALSE}
 p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
@@
 
-```{r,echo=FALSE}
+```{r,echo=FALSE,eval=FALSE}
 p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
@@
 
-```{r,echo=FALSE}
+```{r,echo=FALSE,eval=FALSE}
 xlist = c("Poverty.f","marital.status.f","Primary.Payer.f2","Usual.Occupation.f2","race.f","stage.f","age.f","chemo.f","surgery.f")
@@
 ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+```
 
@@
 
-```{r,echo=FALSE,eval=FALSE}
+```{r,echo=FALSE}
 

### 2020-07-26 manuscript_code_20200722.Rmd
tables

@@
 library(RColorBrewer)
-# library(ggpubr)
+library(ggpubr)
 
@@
 dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
-dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
+# dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
 cat(paste("Three outliers excluded"),"\n")
@@
 ```{r,echo=FALSE,eval=FALSE}
-xlist = c("Poverty.f","marital.status.f","Primary.Payer.f2","Usual.Occupation.f2","race.f","stage.f","age.f","chemo.f","surgery.f")
+xlist = explvars[4:length(explvars)]
 
 x = xlist[1]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+p <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                color = x, palette="jco")
@@
 x = xlist[2]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
                color = x, palette="jco")
@@
 x = xlist[3]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
+p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
                color = x, palette="jco")
@@
 ## Figure code - boxplot
-
-x = "Poverty"
-tmp = cdmd.pvt[,c("Pain","Poverty.f")]
-colnames(tmp) = c("Pain",x)
-tmp$Poverty = factor(tmp$Poverty)
-tmp$Poverty = relevel(tmp$Poverty,"low")
-base <- ggboxplot(tmp, x=x, y="Pain",
+xlist = explvars[4:length(explvars)]
+
+# x = "Poverty"
+# tmp = cdmd.pvt[,c("Pain","Poverty.f")]
+# colnames(tmp) = c("Pain",x)
+# tmp$Poverty = factor(tmp$Poverty)
+# tmp$Poverty = relevel(tmp$Poverty,"low")
+# base <- ggboxplot(tmp, x=x, y="Pain",
+#                   color = x, palette="jco",legend="none") +
+#   labs(x = NULL, y = "Pain", title = x) + 
+#   font("title",size=20,color="black") + 
+#   font("x.text", size = 20, color = "black") +
+#   font("ylab", size = 20, color = "black") + 
+#   font("y.text", size = 20, color = "black") +
+#   stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
+#                      label.y = 10.5)
+# p <- base +
+#   theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
+# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = xlist[2]
+base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                   color = x, palette="jco",legend="none") +
@@
   theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
-ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
-
-x = "Marital.Status"
-tmp <-
-  cdmd.pvt %>%
-  filter(marital.status.f2 %in% c("Married","Single")) %>%
-  select("Pain","marital.status.f2")
-colnames(tmp) = c("Pain",x)
-class(tmp[,2])
-base <- ggboxplot(tmp, x=x, y="Pain",
+p
+# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+
+x = xlist[3]
+base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                   color = x, palette="jco",legend="none") +
@@
   stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
-                     label.y = 10.5)
+                     label.y=10.5)
 p <- base +
   theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
-ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
-
-x = "Primary.Payer"
-tmp <-
-  cdmd.pvt %>%
-  filter(Primary.Payer.f2 %in% c("Insurance","Medicaid")) %>%
-  select("Pain","Primary.Payer.f2")
-
-colnames(tmp) = c("Pain",x)
-class(tmp[,2])
-
-base <- ggboxplot(tmp, x=x, y="Pain",
-                  color = x, palette="jco",legend="none") +
-  labs(x = NULL, y = "Pain", title = x) + 
-  font("title",size=20,color="black") + 
-  font("x.text", size = 20, color = "black") +
-  font("ylab", size = 20, color = "black") + 
-  font("y.text", size = 20, color = "black") +
-  stat_compare_means(size=7, aes(label = "p.format"),show.legend=F,hide.ns=T,
-                     label.y = 10.5)
-p <- base +
-  theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
-
-ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+p
+# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
 
 x = "Occupation"
-tmp <-
-  cdmd.pvt %>%
-  filter(Usual.Occupation.f2 %in% c("Employed","Retired","Unemployed")) %>%
-  select("Pain","Usual.Occupation.f2")
-
-colnames(tmp) = c("Pain",x)
-class(tmp[,2])
-
 my_comparisons <- list( c("Employed", "Retired"), c("Retired", "Unemployed"), c("Employed", "Unemployed") )
-base <- ggboxplot(tmp, x=x, y="Pain",
+base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
                   color = x, palette="jco",legend="none") +
@@
   stat_compare_means(comparisons = my_comparisons, aes(label = ..p.signif..),hide.ns=T)
-
 p <- base +
   theme(plot.title=element_text(face="bold",size=20,margin=margin(t=10,b=20)))
-ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
+p
+# ggsave(p, file=paste0(x,".pdf"),width=6, height=6)
 

### 2020-07-28 manuscript_code_20200722.Rmd
blight added

@@
 
+transform.p = function(x,digits=3) {
+  y = as.character(round(x,digits=digits))
+  y[which(y==0)] = "<.001"
+  return(y)
+}
+
 ```
@@
 
-unknownRaceN = length(which(dmd$race.f %in% c("Other","Unknown")))
-dmd = dmd[-which(dmd$race.f %in% c("Other","Unknown")),]
-cat(paste("Unknown and Other in Race removed (",unknownRaceN,"), Black and White remained N =",dim(dmd)[1]),"\n")
+# unknownRaceN = length(which(dmd$race.f %in% c("Other","Unknown")))
+# dmd = dmd[-which(dmd$race.f %in% c("Other","Unknown")),]
+dmd$race.f[which(dmd$race.f %in% c("Other","Unknown"))] = "Unknown"
+# cat(paste("Unknown and Other in Race removed (",unknownRaceN,"), Black and White remained N =",dim(dmd)[1]),"\n")
 dmd$race.f = factor(dmd$race.f)
@@
 dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
-# dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
+dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
 cat(paste("Three outliers excluded"),"\n")
@@
          "Race","Age",
-         "Histology",
+         "Metastasis","Histology",
          "Primary.Site.1","Primary.Site.2",
-         "HER2","ER","PR","Metastasis")
+         "HER2","ER","PR")
 ```
@@
              "Race","Age",
-             "Histology",
+             "Metastasis","Histology",
              "Primary.Site.1","Primary.Site.2",
-             "HER2","ER","PR","Metastasis")
+             "HER2","ER","PR")
 
@@
   cname = colnames(final.data)[j]
+  formula = as.formula(paste("Pain ~",cname))
   if (class(x)=="numeric") {
+    if (! cname == "Pain") {
+      fit = glm(formula, data = final.data, family = "poisson")
+      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
+    } else {
       tmp.sum = t(t(c(NA,NA)))
+    }
   } else {
@@
   cname = colnames(final.data)[j]
+  formula = as.formula(paste("Pain ~",cname))
   if (class(x)=="numeric") {
-    tmp.sum = t(t(c(1,NA)))
+    if (! cname == "Pain") {
+      fit = glm(formula, data = final.data, family = "poisson")
+      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
+    } else {
+      tmp.sum = t(t(c(NA,NA)))
+    }
   } else {
     tmp.sum = t(t(rep(NA,length(levels(x)))))
-    tmp.sum[1,1] = 1
+    pval = compare_means(formula,data=final.data)
+    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
   }
@@
 
-compare_means(Pain ~ Occupation,data=final.data)
+pval = compare_means(Pain ~ Occupation,data=final.data)
+min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))])
 ```
@@
 
-x = xlist[2]
+x = xlist[4]
 base <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
@@
                "Race","Age",
-               "Histology",
+               "Metastasis","Histology",
                "Primary.Site.1","Primary.Site.2",
-               "HER2","ER","PR","Metastasis")
+               "HER2","ER","PR")
 
@@
 
-transform.p = function(x) {
-  y = as.character(x)
-  y[which(y==0)] = "<.001"
-  return(y)
-}
 final.table$Unadj.P = transform.p(final.table$Unadj.P)
@@
   add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
-  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.count)))) %>%
-  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.count)))) %>%
-  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.count)))) %>%
-  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
-  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
-  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.count)))) %>%
-  add_row(,.before=14+which(grepl(pattern="Metastasis",rownames(table.count))))
+  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
+  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.count)))) %>%
+  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.count)))) %>%
+  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.count)))) %>%
+  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
+  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
+  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.count))))
 colnames(table.count) = c("Estimate","95% CI","P-value",
@@
   add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
-  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
-  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.zero)))) %>%
-  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
-  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
-  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
-  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.zero)))) %>%
-  add_row(,.before=14+which(grepl(pattern="Metastasis",rownames(table.zero))))
+  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
+  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
+  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.zero)))) %>%
+  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
+  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
+  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
+  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.zero))))
 colnames(table.zero) = c("Estimate","95% CI","P-value",

### 2020-07-28 manuscript_code_20200728.Rmd
Blight added

--- 2020-07-28 manuscript_code_20200722.Rmd
@@
 colnames(patients_file)[which(colnames(patients_file)=="POVERTY")] = "Poverty"
-colnames(patients_file)[which(colnames(patients_file)=="VACANCY")] = "Vacancy"
+colnames(patients_file)[which(colnames(patients_file)=="VACANCY")] = "Blight"
 colnames(patients_file)[which(colnames(patients_file)=="WALKABILITY")] = "Walkability"
@@
 
-* "joined_zipcode" file merged to get data about poverty, vacancy, unemployment levels for more zip codes.   
+* "joined_zipcode" file merged to get data about poverty, Blight, unemployment levels for more zip codes.   
 * Poverty data (from "joined_zipcode")
@@
 dmd.pvt = dmd.pvt[-which(dmd.pvt$Poverty>70),]
-dmd.pvt = dmd.pvt[-which((dmd.pvt$Vacancy>40) & (dmd.pvt$mean.Pain<2)),]
+dmd.pvt = dmd.pvt[-which((dmd.pvt$Blight>40) & (dmd.pvt$mean.Pain<2)),]
 cat(paste("Three outliers excluded"),"\n")
@@
   select("Pain",
-         "Poverty","Vacancy",
+         "Poverty","Blight",
          "Occupation","Marital.status","Insurance.type",
@@
 respvar = "Pain"
-explvars = c("Pain","Poverty","Vacancy",
+explvars = c("Pain","Poverty","Blight",
              "Occupation","Marital.status","Insurance.type",
@@
 ```{r,echo=FALSE}
-explvars = c("Pain","Poverty","Vacancy",
+explvars = c("Pain","Poverty","Blight",
              "Occupation","Marital.status","Insurance.type",
@@
 
-**_Vacancy variable_**:
+**_Blight variable_**:
 
@@
 
-hist(cdmd.pvt$Vacancy,main="Vacancy")
+hist(cdmd.pvt$Blight,main="Blight")
 cat(paste("Number of zipcode =",length(unique(cdmd.pvt$zip))),"\n")
-cat(paste("Median of Vacancy =",round(median(cdmd.pvt$Vacancy),digits=3)),"\n")
+cat(paste("Median of Blight =",round(median(cdmd.pvt$Blight),digits=3)),"\n")
 
@@
 
-### Vacancy
-
-* Pain ~ Vacancy
+### Blight
+
+* Pain ~ Blight
 
 ```{r,echo=FALSE,eval=FALSE}
-p.glm = glm(Pain ~ Vacancy, data = final.data, family = "poisson")
+p.glm = glm(Pain ~ Blight, data = final.data, family = "poisson")
 # summary(p.glm)
@@
 
-meandata = data.frame(Vacancy = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
+meandata = data.frame(Blight = aggregate(Blight ~ zip,cdmd.pvt,mean)$Blight,
                       Pain = aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
-                      zip = aggregate(Vacancy ~ zip,cdmd.pvt,mean)$zip)
+                      zip = aggregate(Blight ~ zip,cdmd.pvt,mean)$zip)
 meandata$size = sapply(meandata$zip,function(x) length(which(cdmd.pvt$zip==x)))
 # hist(round(meandata$Pain))
-# summary(glm(round(meandata$Pain) ~ meandata$Vacancy,family = "poisson"))
-# summary(glm.nb(round(meandata$Pain) ~ meandata$Vacancy))
+# summary(glm(round(meandata$Pain) ~ meandata$Blight,family = "poisson"))
+# summary(glm.nb(round(meandata$Pain) ~ meandata$Blight))
 
@@
 
-pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
+pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
 par(mar=c(4,4,2,2))
-plot(x=meandata$Vacancy,y=meandata$Pain,
+plot(x=meandata$Blight,y=meandata$Pain,
      xlim=c(0,60),ylim=c(0,6.5),col=brewer.pal(n=11,"BrBG")[9],
@@
 axis(side=2, tck=-0.015,lwd=0,line=-.8,cex.axis=1.5)
-mtext(side=1, "Vacancy", line=2.5, cex=1.5)
+mtext(side=1, "Blight", line=2.5, cex=1.5)
 mtext(side=2, "Pain average", line=2, cex=1.5)
@@
 ```{r,echo=FALSE,eval=FALSE}
-p.glm = glm(Pain ~ Vacancy, data = cdmd.pvt, family = "poisson")
+p.glm = glm(Pain ~ Blight, data = cdmd.pvt, family = "poisson")
 # summary(p.glm)
@@
 cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
-# pdf(6,6,file=paste("Pain_Vacancy_plot.pdf"))
-plot(x=meandata$Vacancy,y=meandata$Pain,cex=cex,xlab="Vacancy",ylab="Pain average")
-# plot(x=aggregate(Vacancy ~ zip,cdmd.pvt,mean)$Vacancy,
+# pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
+plot(x=meandata$Blight,y=meandata$Pain,cex=cex,xlab="Blight",ylab="Pain average")
+# plot(x=aggregate(Blight ~ zip,cdmd.pvt,mean)$Blight,
 #      y=aggregate(Pain ~ zip,cdmd.pvt,mean)$Pain,
-#      xlab="Vacancy",ylab=paste(pain.type,"Pain mean within Zip"))
+#      xlab="Blight",ylab=paste(pain.type,"Pain mean within Zip"))
 x=seq(0,80,by=1)
@@
                "Race","Age")
-explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
+explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
                "Cancer.stage","Chemo","Surgery",
                "Race","Age")
-explvars = c("Poverty","Vacancy",
+explvars = c("Poverty","Blight",
              "Occupation","Marital.status","Insurance.type",
@@
 vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
-tablevars = c("(Intercept)","Poverty","Vacancy",unlist(vars.list)) # all variables that will be printed in the table
+tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table
 
@@
 
-explvars.2 = c("Vacancy","Occupation","Marital.status","Insurance.type",
+explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
                "Cancer.stage","Chemo","Surgery",
@@
 
-explvars = c("Poverty","Vacancy",
+explvars = c("Poverty","Blight",
              "Occupation","Marital.status","Insurance.type",

### 2020-07-28 R/resultsTable.R

@@ -1,0 +1,147 @@
+#'
+#' @export
+get_resultsTable_base = function(formula,data) {
+  require(sjPlot)
+  formula.bin = as.formula(paste(sub("Pain",replacement="bin.Pain",as.character(formula))[c(2,3)],collapse=" ~ "))
+  m.slr = lm(formula, data = data)
+  m.glm.nb = glm.nb(formula, data = data)
+  m.glm.bin = glm(formula.bin, data = data, family="binomial")
+  m.zinb = zeroinfl(formula, data = data, dist="negbin")
+  m.hurdle = hurdle(formula, data = data,dist="negbin")
+
+  tab_model(m.slr,m.glm.nb,m.glm.bin,m.zinb,m.hurdle, digits = 3,
+            dv.labels = c("SLR","GLM-NB","GLM-BIN","ZINB","Hurdle"),
+            title = "Results Table",
+            transform="exp",
+            show.zeroinf=TRUE, show.intercept=TRUE, show.se=FALSE, show.ci=FALSE, collapse.se=FALSE,
+            show.p = TRUE, p.style="both")
+}
+
+#'
+#' @export
+get_resultsTable4painTypes = function(formula,data,model=c("zeroinfl.nb","zeroinfl.poi","lm","glm.nb","glm.poi","hurdle.nb","hurdle.poi")) {
+  require(sjPlot)
+  require(pscl)
+
+  model = match.arg(model,c("zeroinfl.nb","zeroinfl.poi","lm","glm.nb","glm.poi","hurdle.nb","hurdle.poi"))
+  painTypes = c("mean","Q1st","Q2nd","Q3rd","max")
+  result = as.list(NULL)
+  for (i in seq_along(painTypes)) {
+    pain.type = painTypes[i]
+    newdata = data
+    colnames(newdata)[which(grepl(pattern=pain.type,colnames(newdata))==T)] = c("Pain","Pain.bfchemo","Pain.afchemo")
+    if (model=="zeroinfl.nb") {
+      result[[i]] = zeroinfl(formula, data = newdata, dist="negbin")
+    } else if (model=="zeroinfl.poi") {
+      result[[i]] = zeroinfl(formula, data = newdata, dist="poisson")
+    } else if (model=="lm") {
+      result[[i]] = lm(formula, data = newdata)
+    } else if (model=="glm.nb") {
+      result[[i]] = glm.nb(formula, data = newdata)
+    } else if (model=="glm.poi") {
+      result[[i]] = glm(formula, data = newdata, family = "poisson")
[…]

### 2020-08-05 manuscript_code_20200728.Rmd
figures updated

@@
 
-## Table 1
+## Table 1 (Base)
 
@@
 
-```{r,echo=FALSE,eval=FALSE}
-respvar = "Pain"
+```{r,echo=FALSE}
 explvars = c("Pain","Poverty","Blight",
@@
              "Cancer.stage","Chemo","Surgery",
-             "Race","Age",
-             "Metastasis","Histology",
-             "Primary.Site.1","Primary.Site.2",
-             "HER2","ER","PR")
-
-table1(as.formula(paste(" ~ ",paste(c(respvar,explvars),collapse=" + "))), data=final.data)
-```
-
-```{r,echo=FALSE}
-explvars = c("Pain","Poverty","Blight",
-             "Occupation","Marital.status","Insurance.type",
-             "Cancer.stage","Chemo","Surgery",
-             "Race","Age",
-             "Metastasis","Histology",
-             "Primary.Site.1","Primary.Site.2",
-             "HER2","ER","PR")
+             "Race","Age")
 
@@
 
+table.sum1 = cbind(do.call("rbind",sapply(1:length(explvars),get_summary)),
+                   do.call("rbind",sapply(1:length(explvars),get_mean)),
+                   do.call("rbind",sapply(1:length(explvars),get_pval)))
+colnames(table.sum1) = c("No. (%)","Pain, mean [Q1,Q3]","P-value")
+
+## htmlTable
+library(htmlTable)
+
+rowlabels = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[1])
+htmlTable(table.sum1,
+          rgroup = c(gsub(pattern="[.]",replacement=" ",x=explvars[1:length(explvars)])),
+          rnames = sapply(rownames(table.sum1),function(x) unlist(strsplit(x,split="_"))[2]),
+          n.rgroup = sapply(explvars, function(x) length(which(rowlabels==x))),
+          caption = "Table 1. Baseline characteristics, mean and interquartile range of pain intensity, nonparametric rank test")
+
+```
+
+## Table 1 (Base + more biomarker variables)
+
+```{r,echo=FALSE,eval=FALSE}
+respvar = "Pain"
+explvars = c("Pain","Poverty","Blight",
+             "Occupation","Marital.status","Insurance.type",
+             "Cancer.stage","Chemo","Surgery",
+             "Race","Age",
+             "Metastasis","Histology",
+             "Primary.Site.1","Primary.Site.2",
+             "HER2","ER","PR")
+
+table1(as.formula(paste(" ~ ",paste(c(respvar,explvars),collapse=" + "))), data=final.data)
+```
+
+```{r,echo=FALSE}
+explvars = c("Pain","Poverty","Blight",
+             "Occupation","Marital.status","Insurance.type",
+             "Cancer.stage","Chemo","Surgery",
+             "Race","Age",
+             "Metastasis","Histology",
+             "Primary.Site.1","Primary.Site.2",
+             "HER2","ER","PR")
+
+get_summary = function(j) {
+  x = final.data[,j]
+  cname = colnames(final.data)[j]
+  if (class(x)=="numeric") {
+    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
+                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
+    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
+  } else {
+    tmp.sum = t(t(summary(x)))
+    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
+    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
+  }
+  return(tmp.sum)
+}
+table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
+
+get_mean = function(j) {
+  x = final.data[,j]
+  cname = colnames(final.data)[j]
+  formula = as.formula(paste("Pain ~",cname))
+  if (class(x)=="numeric") {
+    if (! cname == "Pain") {
+      fit = glm(formula, data = final.data, family = "poisson")
+      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
+    } else {
+      tmp.sum = t(t(c(NA,NA)))
+    }
+  } else {
+    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
+  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
+  }
+  return(tmp.sum)
+}
+
+get_pval = function(j) {
+  x = final.data[,j]
+  cname = colnames(final.data)[j]
+  formula = as.formula(paste("Pain ~",cname))
+  if (class(x)=="numeric") {
+    if (! cname == "Pain") {
+      fit = glm(formula, data = final.data, family = "poisson")
+      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
+    } else {
+      tmp.sum = t(t(c(NA,NA)))
+    }
+  } else {
+    tmp.sum = t(t(rep(NA,length(levels(x)))))
+    pval = compare_means(formula,data=final.data)
+    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
+  }
+  return(tmp.sum)
+}
+
 table.sum1 = cbind(do.call("rbind",sapply(1:ncol(final.data),get_summary)),
@@
 
-## Figure 2
+## Scatter plots
 
@@
 cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
+
+x=seq(0,70,by=1)
+yy.pred = predict(p.glm, data.frame(Poverty=x), type="link", se=T)
+yy = exp(yy.pred$fit)
+yy.upper = exp(yy.pred$fit + 1.96*yy.pred$se)
+yy.lower = exp(yy.pred$fit - 1.96*yy.pred$se)
 
@@
 #        cex=cex,lwd=1.5)
-x=seq(0,70,by=1)
-points(x=x,y=fhat(x),type="l",lwd=3,
+points(x=x,y=yy,type="l",lwd=3,
        col=brewer.pal(n=9,"YlOrRd")[7])
+polygon(c(x,rev(x)), c(yy.upper,rev(yy.lower)), col="#FF000050", border=NA)
+
+legend("bottomright",bty="n",cex=1.5,
+       legend=c(paste("exp(coef) =",round(exp(p.glm$coefficients[2]),digits=2)),
+                paste("(P-value < .001)"),
+                paste("")))
+
 dev.off()
@@
 cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
+
+x=seq(0,60,by=1)
+yy.pred = predict(p.glm, data.frame(Blight=x), type="link", se=T)
+yy = exp(yy.pred$fit)
+yy.upper = exp(yy.pred$fit + 1.96*yy.pred$se)
+yy.lower = exp(yy.pred$fit - 1.96*yy.pred$se)
 
@@
 
-x=seq(0,53,by=1)
-points(x=x,y=fhat(x),type="l",lwd=3,
+points(x=x,y=yy,type="l",lwd=3,
        col=brewer.pal(n=9,"YlOrRd")[7])
+polygon(c(x,rev(x)), c(yy.upper,rev(yy.lower)), col="#FF000050", border=NA)
+
+legend("bottomright",bty="n",cex=1.5,
+       legend=c(paste("exp(coef) =",round(exp(p.glm$coefficients[2]),digits=2)),
+                paste("(P-value < .001)"),
+                paste("")))
 dev.off()

### 2020-08-09 manuscript_code_20200809.Rmd
table updated

--- 2020-08-05 manuscript_code_20200728.Rmd
@@
 colnames(dmd)[which(colnames(dmd)=="Tobacco.History")] = "Tobacco.history"
+colnames(dmd)[which(colnames(dmd)=="count.Pain")] = "Visits"
 ```
@@
          "Occupation","Marital.status","Insurance.type",
-         "Cancer.stage","Chemo","Surgery",
+         "Cancer.stage","Chemo","Visits",
          "Race","Age",
@@
 
-# Summary of variables
-
-## Table 1 (Base)
+# Table1: Summary of variables
+
+## Base
 
@@
              "Occupation","Marital.status","Insurance.type",
-             "Cancer.stage","Chemo","Surgery",
+             "Cancer.stage","Chemo","Visits",
              "Race","Age")
@@
   cname = colnames(final.data)[j]
-  if (class(x)=="numeric") {
+  if (class(x) %in% c("numeric","integer")) {
     tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
@@
   formula = as.formula(paste("Pain ~",cname))
-  if (class(x)=="numeric") {
+  if (class(x) %in% c("numeric","integer")) {
     if (! cname == "Pain") {
@@
 
-## Table 1 (Base + more biomarker variables)
+## Base + more biomarker variables
 
@@
              "Occupation","Marital.status","Insurance.type",
-             "Cancer.stage","Chemo","Surgery",
+             "Cancer.stage","Chemo","Visits",
              "Race","Age",
@@
 
-get_summary = function(j) {
-  x = final.data[,j]
-  cname = colnames(final.data)[j]
-  if (class(x)=="numeric") {
-    tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
-                    paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
-    rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
-  } else {
-    tmp.sum = t(t(summary(x)))
-    tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
-    rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
-  }
-  return(tmp.sum)
-}
-table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
-
-get_mean = function(j) {
-  x = final.data[,j]
-  cname = colnames(final.data)[j]
-  formula = as.formula(paste("Pain ~",cname))
-  if (class(x)=="numeric") {
-    if (! cname == "Pain") {
-      fit = glm(formula, data = final.data, family = "poisson")
-      tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
-    } else {
-      tmp.sum = t(t(c(NA,NA)))
-    }
-  } else {
-    tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
-  paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
-  }
-  return(tmp.sum)
-}
-
-get_pval = function(j) {
-  x = final.data[,j]
-  cname = colnames(final.data)[j]
-  formula = as.formula(paste("Pain ~",cname))
-  if (class(x)=="numeric") {
-    if (! cname == "Pain") {
-      fit = glm(formula, data = final.data, family = "poisson")
-      tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
-    } else {
-      tmp.sum = t(t(c(NA,NA)))
-    }
-  } else {
-    tmp.sum = t(t(rep(NA,length(levels(x)))))
-    pval = compare_means(formula,data=final.data)
-    tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
-  }
-  return(tmp.sum)
-}
+# get_summary = function(j) {
+#   x = final.data[,j]
+#   cname = colnames(final.data)[j]
+#   if (class(x)=="numeric") {
+#     tmp.sum = t(t(c(paste0(round(mean(x),digits=2)," (",round(sd(x),digits=2),")"),
+#                     paste0(round(median(x),digits=2)," [",round(min(x),digits=2),",",round(max(x),digits=2),"]"))))
+#     rownames(tmp.sum) = paste(cname,c("Mean (SD)","Median [Min,Max]"),sep="_")
+#   } else {
+#     tmp.sum = t(t(summary(x)))
+#     tmp.sum[,1] = paste(tmp.sum[,1],paste0("(",round(100*(tmp.sum[,1]/sum(tmp.sum[,1])),digits=1),"%)"))
+#     rownames(tmp.sum) = paste(cname,rownames(tmp.sum),sep="_")
+#   }
+#   return(tmp.sum)
+# }
+# table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
+# 
+# get_mean = function(j) {
+#   x = final.data[,j]
+#   cname = colnames(final.data)[j]
+#   formula = as.formula(paste("Pain ~",cname))
+#   if (class(x)=="numeric") {
+#     if (! cname == "Pain") {
+#       fit = glm(formula, data = final.data, family = "poisson")
+#       tmp.sum = t(t(c(round(exp(coef(summary(fit))[2,1]),digits=2),NA)))
+#     } else {
+#       tmp.sum = t(t(c(NA,NA)))
+#     }
+#   } else {
+#     tmp.sum = t(t(sapply(split(final.data[,"Pain"],final.data[,j]),function(x) 
+#   paste0(round(mean(x),digits=2)," [",paste0(round(quantile(x,c(0.25,0.75)),digits=2),collapse=","),"]"))))
+#   }
+#   return(tmp.sum)
+# }
+# 
+# get_pval = function(j) {
+#   x = final.data[,j]
+#   cname = colnames(final.data)[j]
+#   formula = as.formula(paste("Pain ~",cname))
+#   if (class(x)=="numeric") {
+#     if (! cname == "Pain") {
+#       fit = glm(formula, data = final.data, family = "poisson")
+#       tmp.sum = t(t(c(transform.p(coef(summary(fit))[2,4],digits=3),NA)))
+#     } else {
+#       tmp.sum = t(t(c(NA,NA)))
+#     }
+#   } else {
+#     tmp.sum = t(t(rep(NA,length(levels(x)))))
+#     pval = compare_means(formula,data=final.data)
+#     tmp.sum[1,1] = transform.p(min(pval$p[which((! pval$group1 %in% "Unknown") & (! pval$group2 %in% "Unknown"))]),digits=3)
+#   }
+#   return(tmp.sum)
+# }
 
@@
 
-## Scatter plots
+# Plots
 
@@
 
-### Poverty
+## Poverty
 
@@
 
-```{r,echo=FALSE,eval=FALSE}
+```{r,echo=FALSE}
 p.glm = glm(Pain ~ Poverty, data = final.data, family = "poisson")
@@
 
-# par(mfrow=c(1,2))
+par(mfrow=c(1,2))
 cex = exp(2*(meandata$size/(max(meandata$size) - min(meandata$size) + 1)))
@@
 
-pdf(6,6,file=paste("Pain_poverty_plot.pdf"))
+# pdf(6,6,file=paste("Pain_poverty_plot.pdf"))
 par(mar=c(4,4,2,2))
@@
 
-dev.off()
+# dev.off()
 
 # par(mfrow=c(3,3))
-pdf(6,6,file=paste("Pain_poverty_legend.pdf"))
-par(mar=c(4,4,2,2))
-plot(x=c(0,1),y=c(0,1),axes=F,cex=0,xlab=NA,ylab=NA)
-# cbind(meandata$size,cex)
-
-x = c(20,40,60,80)
-y = c(1.5,2.5,4,6)
-points(x=rep(0.4,length(x)),
-       y=seq(0.8,0.4,length.out=length(x)),
-       cex=y,lwd=2.5,col=brewer.pal(n=11,"BrBG")[9])
-text(x=rep(0.55,length(x)),
-     y=seq(0.8,0.4,length.out=length(x)),
-     labels=x,cex=1.5)
-text(x=0.5,y=0.95,
-     labels="Number of patients",cex=1.5)
-dev.off()
-
-```
-
-### Blight
+# pdf(6,6,file=paste("Pain_poverty_legend.pdf"))
+# par(mar=c(4,4,2,2))
+# plot(x=c(0,1),y=c(0,1),axes=F,cex=0,xlab=NA,ylab=NA)
+# # cbind(meandata$size,cex)
+# 
+# x = c(20,40,60,80)
+# y = c(1.5,2.5,4,6)
+# points(x=rep(0.4,length(x)),
+#        y=seq(0.8,0.4,length.out=length(x)),
+#        cex=y,lwd=2.5,col=brewer.pal(n=11,"BrBG")[9])
+# text(x=rep(0.55,length(x)),
+#      y=seq(0.8,0.4,length.out=length(x)),
+#      labels=x,cex=1.5)
+# text(x=0.5,y=0.95,
+#      labels="Number of patients",cex=1.5)
+# dev.off()
+
+```
+
+## Blight
 
@@
 
-```{r,echo=FALSE,eval=FALSE}
+```{r,echo=FALSE}
 p.glm = glm(Pain ~ Blight, data = final.data, family = "poisson")
@@
 
-pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
+# pdf(6,6,file=paste("Pain_Blight_plot.pdf"))
 par(mar=c(4,4,2,2))
@@
                 paste("")))
-dev.off()
+# dev.off()
 
@@
 
-# Analysis
-
-## Figure 2
-
-# Significance
+# Table 2: Regression model
+
+## Base
 
@@
 
-## Table 2
-
 ```{r,echo=FALSE}
@@
 explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Surgery",
+               "Cancer.stage","Chemo","Visits",
                "Race","Age")
 explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Surgery",
+               "Cancer.stage","Chemo","Visits",
                "Race","Age")
@@
 
-## Poisson count model
+### Poisson count model
 ```{r,echo=FALSE}
@@
 # library(Gmisc)
-vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
-vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
-tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table
+
+vars.list = sapply(explvars,function(x) {
+  x.level <- levels(final.data[,which(colnames(final.data)==x)])
+  if (! is.null(x.level)) {
+    x.level[1] <- paste(x.level[1],"(reference)")
+    return(x.level)
+  } else {
+    return(x)
+  }
+})
+tablevars = c("(Intercept)",unlist(vars.list)) # all variables that will be printed in the table
 
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
-  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
+  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.count)))) %>%
   add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
@@
 htmlTable(table.count,
-          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
+          rgroup = c("",gsub(pattern="[.]",replacement=" ",x=explvars)),
           rnames = tablevars,
-          n.rgroup = c(1,1,1,sapply(vars.list,length)),
+          n.rgroup = c(1,sapply(vars.list,length)),
           cgroup = c("Unadjusted","Adjusted","Adjusted"),
@@
 
-## Inflation model
+### Inflation model
 ```{r,echo=FALSE}
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
-  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
+  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
   add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
@@
 htmlTable(table.zero,
-          rgroup = c("","","",gsub(pattern="[.]",replacement=" ",x=explvars[3:length(explvars)])),
+          rgroup = c("",gsub(pattern="[.]",replacement=" ",x=explvars)),
           rnames = tablevars,
-          n.rgroup = c(1,1,1,sapply(vars.list,length)),
+          n.rgroup = c(1,sapply(vars.list,length)),
           cgroup = c("Unadjusted","Adjusted","Adjusted"),
@@
 
-# Basic model + more biomarker variables
+## Base + more biomarker variables
 
@@
 explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Surgery",
+               "Cancer.stage","Chemo","Visits",
                "Race","Age",
@@
 explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Surgery",
+               "Cancer.stage","Chemo","Visits",
                "Race","Age",
@@
 
-## Poisson count model:
+### Poisson count model:
 ```{r,echo=FALSE}
@@
 # library(Gmisc)
-vars.list = sapply(explvars[-c(1,2)],function(x) levels(final.data[,which(colnames(final.data)==x)]))
-vars.list = lapply(vars.list,function(x) {x[1]<-paste(x[1],"(reference)"); return(x)})
-tablevars = c("(Intercept)","Poverty","Blight",unlist(vars.list)) # all variables that will be printed in the table
+vars.list = sapply(explvars,function(x) {
+  x.level <- levels(final.data[,which(colnames(final.data)==x)])
+  if (! is.null(x.level)) {
+    x.level[1] <- paste(x.level[1],"(reference)")
+    return(x.level)
+  } else {
+    return(x)
+  }
+})
+tablevars = c("(Intercept)",unlist(vars.list)) # all variables that will be printed in the table
 
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
-  add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
+  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
   add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
@@
 
-## Inflation model:
+### Inflation model:
 ```{r,echo=FALSE}

### 2020-08-10 manuscript_code_20200809.Rmd
table updated

@@
          "Occupation","Marital.status","Insurance.type",
-         "Cancer.stage","Chemo","Visits",
+         "Cancer.stage","Chemo",
          "Race","Age",
-         "Metastasis","Histology",
+         "Visits","Metastasis","Histology",
          "Primary.Site.1","Primary.Site.2",
@@
              "Occupation","Marital.status","Insurance.type",
-             "Cancer.stage","Chemo","Visits",
+             "Cancer.stage","Chemo",
              "Race","Age")
@@
 }
-table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
+# table.sum1 = do.call("rbind",sapply(1:ncol(final.data),get_summary))
 
@@
              "Occupation","Marital.status","Insurance.type",
-             "Cancer.stage","Chemo","Visits",
+             "Cancer.stage","Chemo",
              "Race","Age",
-             "Metastasis","Histology",
+             "Histology",
              "Primary.Site.1","Primary.Site.2",
@@
              "Occupation","Marital.status","Insurance.type",
-             "Cancer.stage","Chemo","Visits",
+             "Cancer.stage","Chemo","Visits","Metastasis",
              "Race","Age",
-             "Metastasis","Histology",
+             "Histology",
              "Primary.Site.1","Primary.Site.2",
@@
 ```{r,echo=FALSE,eval=FALSE}
-xlist = explvars[4:length(explvars)]
-
-x = xlist[1]
-p <- ggboxplot(final.data, x=x, y="Pain",remove="Unknown",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[2]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[3]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",remove="Unknown",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[4]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[5]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[6]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[7]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[8]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-x = xlist[9]
-p <- ggboxplot(cdmd.pvt, x=x, y="Pain",
-               color = x, palette="jco")
-p + stat_compare_means()
-
-```
-
-```{r,echo=FALSE,eval=FALSE}
 ## Figure code - boxplot
@@
 explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Visits",
+               "Cancer.stage","Chemo",
                "Race","Age")
 explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Visits",
+               "Cancer.stage","Chemo",
                "Race","Age")
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
-  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.count)))) %>%
-  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
-  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count))))
+  # add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
+  add_row(,.before=5+which(grepl(pattern="Black",rownames(table.count)))) %>%
+  add_row(,.before=6+which(grepl(pattern="Age",rownames(table.count))))
 colnames(table.count) = c("Estimate","95% CI","P-value",
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
-  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
-  add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
-  add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero))))
+  # add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
+  add_row(,.before=5+which(grepl(pattern="Black",rownames(table.zero)))) %>%
+  add_row(,.before=6+which(grepl(pattern="Age",rownames(table.zero))))
 colnames(table.zero) = c("Estimate","95% CI","P-value",
@@
 
+```{r,echo=FALSE,eval=FALSE}
+# estimated probability and 95% CI:
+
+get_estprob = function(estimates,x) {
+  y = sum(estimates*x)
+  return(exp(y)/(1+exp(y)))
+}
+## poverty using unadjusted model
+tmp.table = final.unadj.list[[1]]
+
+apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,0)))
+
+apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,60)))
+
+## insurance type using adjusted model
+tmp.table = final.adj.1
+tmp.table = tmp.table[grepl(pattern="zero",rownames(tmp.table)),]
+x = c(1,mean(final.data$Poverty),0,0,0,0,0,0,0,0,1,0,1,0)
+cbind(rownames(tmp.table),x)
+
+apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,mean(final.data$Poverty),0,0,0,0,0,0,0,0,0,0,0,0,0,0)))
+
+apply(log(tmp.table[grepl(pattern="zero",rownames(tmp.table)),1:3]),2,function(t) get_estprob(estimates=t,x=c(1,mean(final.data$Poverty),0,0,0,0,0,0,1,0,0,0,0,0,0,0)))
+
+table.zero = final.table[which(grepl("zero",rownames(final.table))),]
+rownames(table.zero) = sapply(rownames(table.zero),function(x) unlist(strsplit(x,split="zero_"))[2]) # extract zero model
+
+coef(final)[grepl(pattern="zero",names(coef(final)))]
+```
+
 ## Base + more biomarker variables
@@
 explvars.1 = c("Poverty","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Visits",
+               "Cancer.stage","Chemo","Visits","Metastasis",
                "Race","Age",
-               "Metastasis","Histology",
+               "Histology",
                "Primary.Site.1","Primary.Site.2",
@@
 explvars.2 = c("Blight","Occupation","Marital.status","Insurance.type",
-               "Cancer.stage","Chemo","Visits",
+               "Cancer.stage","Chemo","Visits","Metastasis",
                "Race","Age",
-               "Metastasis","Histology",
+               "Histology",
                "Primary.Site.1","Primary.Site.2",
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.count)))) %>%
-  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.count)))) %>%
+  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.count)))) %>%
+  add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
   add_row(,.before=6+which(grepl(pattern="Black",rownames(table.count)))) %>%
   add_row(,.before=7+which(grepl(pattern="Age",rownames(table.count)))) %>%
-  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
-  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.count)))) %>%
-  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.count)))) %>%
-  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.count)))) %>%
-  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
-  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
-  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.count))))
+  # add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.count)))) %>%
+  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.count)))) %>%
+  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.count)))) %>%
+  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.count)))) %>%
+  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.count)))) %>%
+  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.count)))) %>%
+  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.count))))
 colnames(table.count) = c("Estimate","95% CI","P-value",
@@
   add_row(,.before=4+which(grepl(pattern="Chemo",rownames(table.zero)))) %>%
-  # add_row(,.before=5+which(grepl(pattern="Surgery",rownames(table.zero)))) %>%
+  # add_row(,.before=5+which(grepl(pattern="Visits",rownames(table.zero)))) %>%
+  add_row(,.before=5+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
   add_row(,.before=6+which(grepl(pattern="Black",rownames(table.zero)))) %>%
   add_row(,.before=7+which(grepl(pattern="Age",rownames(table.zero)))) %>%
-  add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
-  add_row(,.before=9+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
-  add_row(,.before=10+which(grepl(pattern="Central",rownames(table.zero)))) %>%
-  add_row(,.before=11+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
-  add_row(,.before=12+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
-  add_row(,.before=13+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
-  add_row(,.before=14+which(grepl(pattern="PRPositive",rownames(table.zero))))
+  # add_row(,.before=8+which(grepl(pattern="Metastasis",rownames(table.zero)))) %>%
+  add_row(,.before=8+which(grepl(pattern="lobular",rownames(table.zero)))) %>%
+  add_row(,.before=9+which(grepl(pattern="Central",rownames(table.zero)))) %>%
+  add_row(,.before=10+which(grepl(pattern="Midline",rownames(table.zero)))) %>%
+  add_row(,.before=11+which(grepl(pattern="HER2Positive",rownames(table.zero)))) %>%
+  add_row(,.before=12+which(grepl(pattern="ERPositive",rownames(table.zero)))) %>%
+  add_row(,.before=13+which(grepl(pattern="PRPositive",rownames(table.zero))))
 colnames(table.zero) = c("Estimate","95% CI","P-value",

### 2020-10-04 manuscript_code_20200808.Rmd
inflation model reversed

--- 2020-08-09 manuscript_code_20200809.Rmd
@@
 dmd.pvt = dmd.pvt[-which((dmd.pvt$Blight>40) & (dmd.pvt$mean.Pain<2)),]
+dmd.pvt = dmd.pvt[-which(dmd.pvt$Visits>40),]
 cat(paste("Three outliers excluded"),"\n")

### 2020-10-04 manuscript_code_20201004.Rmd
inflation model reversed

--- 2020-08-10 manuscript_code_20200809.Rmd
@@
 rownames(final.table) = table.explvars
+final.table[grepl(pattern="zero",rownames(final.table)),c(1:3,5:7,9:11)] = apply(final.table[grepl(pattern="zero",rownames(final.table)),c(1:3,5:7,9:11)],2,
+                  function(x) {
+                    return(round(1/x,digits=3))
+                  })
 

### 2021-02-18 manuscript_code_20201004.Rmd
total days

--- 2020-08-10 manuscript_code_20200809.Rmd
@@
 rownames(final.table) = table.explvars
+final.table[grepl(pattern="zero",rownames(final.table)),c(1:3,5:7,9:11)] = apply(final.table[grepl(pattern="zero",rownames(final.table)),c(1:3,5:7,9:11)],2,
+                  function(x) {
+                    return(round(1/x,digits=3))
+                  })
+final.table[grepl(pattern="zero",rownames(final.table)),c(2,3)] = final.table[grepl(pattern="zero",rownames(final.table)),c(3,2)]
+
+final.table[grepl(pattern="zero",rownames(final.table)),c(6,7)] = final.table[grepl(pattern="zero",rownames(final.table)),c(7,6)]
+
+final.table[grepl(pattern="zero",rownames(final.table)),c(10,11)] = final.table[grepl(pattern="zero",rownames(final.table)),c(11,10)]
 

### 2021-02-18 R/total_days.R

@@ -1,0 +1,6 @@
+total_days = function(x) {
+  return(max(pcm_file[which(pcm_file$profile_key==x),"SURVEY_DATE"]) - min(pcm_file[which(pcm_file$profile_key==x),"SURVEY_DATE"])+1)
+}
+
+pcm_days = sapply(profile_key,total_days)
+mean(pcm_days)/365
